---
title: "Silo3_9_ASCA_logNSAFs_FunctionalAnalysis"
author: "Shelly Trigg"
date: "9/14/2019"
output: html_document
---

**load libraries**
```{r}
#load libraries
library(plyr)
library(tidyr)
library(GSEABase)
library(reshape2)
library(gplots)
library(viridis)
library(topGO)
```

**read in and format uniprot data**
```{r}
#read in uniprot mapping
uniprot <- read.csv("../../Data/all_giga-uniprot-blastP-out.nopipe.annotations.tab", sep ="\t", header = FALSE, stringsAsFactors = FALSE)
#rename uniprot columns
colnames(uniprot) <- c("protein_ID","Entry", "Entry_name", "perc_ident_match", "align_len", "num_mismatch", "num_gaps","querStart", "querEnd", "subjStart", "subjEnd", "evalue", "bitscore","Entry.1","Entry_name.1", "Protein_names", "Gene_names", "Organism", "Protein_length","Pathway", "GO_bp", "GO","GO_IDs", "Protein_fams")
#convert columns that should be numeric to numeric
str(uniprot)
uniprot[,4:13] <- lapply(uniprot[,4:13], as.numeric)
#NAs introduced by coercion
#this is because the entries that are unmappaed have NA values
str(uniprot)

#filter for choyp proteins and for evalue <= 10x10^-10 (unmapped proteins get removed here)
uniprot_filtered <- uniprot[grep("CHOYP",uniprot$protein_ID),]
uniprot_filtered <- uniprot[which(uniprot$evalue <= 1*10^-10),]


```

**read in and format ASCA logNSAF time and temp sig. data**
```{r}
silo39_time_ASCA <- read.csv("../ASCA/ASCA_TimeProts_avgNSAF.csv", stringsAsFactors = FALSE)
colnames(silo39_time_ASCA)[1] <- "protein_ID"

silo39_temp_ASCA <- read.csv("../ASCA/ASCA_TempProts_avgNSAF.csv", stringsAsFactors = FALSE)
colnames(silo39_temp_ASCA)[1] <- "protein_ID"

```

**read in all proteins detected specific for each comparison to generate GO background 
```{r}
silo39_all_prots_t <- read.csv("../Technical_Replicates_PCA/silo3and9_nozerovals_NSAF_AVGs.csv", stringsAsFactors = FALSE)

#remove sample information and transform the data to get protiens listed as rows instead of columns
silo39_all_prots <- data.frame(t(silo39_all_prots_t[,-c(1:2)]))

#add the protein names as the first column
silo39_all_prots <- cbind(rownames(silo39_all_prots), silo39_all_prots)

#rename column 1 to protein_ID and the rest to the sample names
colnames(silo39_all_prots) <- c("protein_ID",paste(silo39_all_prots_t$dpf, silo39_all_prots_t$temp, sep = "_"))

```

**Match GO terms to sig protein lists**

```{r}
#Get GO IDs for silo 3 and 9 sig. proteins
silo39_sig_pro_GO <- merge(silo39_time_ASCA, uniprot_filtered[,c("protein_ID","GO_IDs")], by ="protein_ID")

#count number of proteins with GO terms
nrow(silo39_sig_pro_GO)

#Since GO IDs are listed all in one column, spread them across multiple columns (1 column /term)
silo39_sig_pro_GOid_term <- data.frame()
for (i in 1:nrow(silo39_sig_pro_GO)){
  silo39_sig_pro_GOid_term_row <- data.frame(t(data.frame(strsplit(as.character(silo39_sig_pro_GO$GO_IDs[i]),'; ', fixed = TRUE))))
  silo39_sig_pro_GOid_term <- rbind.fill(silo39_sig_pro_GOid_term,silo39_sig_pro_GOid_term_row)
}

#add protein IDs back to GO IDs
silo39_sig_pro_GOid_term <- cbind(silo39_sig_pro_GO[,"protein_ID"], silo39_sig_pro_GOid_term)
#this results in a table with protein ID listed in one column and each next column contains a GO ID that was listed with the protein in Uniprot DB.
#convert all factors to  characters
silo39_sig_pro_GOid_term <- data.frame(lapply(silo39_sig_pro_GOid_term, as.character), stringsAsFactors = FALSE)
#proteins that don't have GO IDs?

#reshape data so that all GO ID columns are gathered in one column called "GO" 
STACKED_silo39_sig_pro_GOid_term <- tidyr::gather(silo39_sig_pro_GOid_term,"col","GO", 2:ncol(silo39_sig_pro_GOid_term))
#exlude middle column which just contains the string "gene" in each row
STACKED_silo39_sig_pro_GOid_term <- STACKED_silo39_sig_pro_GOid_term[,-2]
#remove duplicate rows
STACKED_silo39_sig_pro_GOid_term <- unique(STACKED_silo39_sig_pro_GOid_term)
colnames(STACKED_silo39_sig_pro_GOid_term)[1] <- "protein_ID"
#remove any rows where GO column has NA value. 
STACKED_silo39_sig_pro_GOid_term <- STACKED_silo39_sig_pro_GOid_term[which(!is.na(STACKED_silo39_sig_pro_GOid_term$GO)),]
#this resulting data frame has two columns "gene" and "GO"
#write.csv(STACKED_sig0.1_pro_GOid_term, "~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/SameDayFCtoTemp/ChiSq_ASCA_clust/intermediate_files/STACKED_sig0.1_pro_GOid_term.csv", quote = FALSE, row.names = FALSE)
```

**MATCH GO TERMS TO BACKGROUND LIST**
This step takes ~10min
```{r}
#Get GO IDs for silo 3 and 9 ALL. proteins
silo39_ALL_pro_GO <-uniprot_filtered[which(uniprot_filtered$protein_ID %in% silo39_all_prots$protein_ID),c("protein_ID","GO_IDs")]

#count number of proteins with GO terms
nrow(silo39_ALL_pro_GO)

#Since GO IDs are listed all in one column, spread them across multiple columns (1 column /term)
silo39_ALL_pro_GOid_term <- data.frame()
for (i in 1:nrow(silo39_ALL_pro_GO)){
  silo39_ALL_pro_GOid_term_row <- data.frame(t(data.frame(strsplit(as.character(silo39_ALL_pro_GO$GO_IDs[i]),'; ', fixed = TRUE))))
  silo39_ALL_pro_GOid_term <- rbind.fill(silo39_ALL_pro_GOid_term,silo39_ALL_pro_GOid_term_row)
}

#add protein IDs back to GO IDs
silo39_ALL_pro_GOid_term <- cbind(silo39_ALL_pro_GO[,"protein_ID"], silo39_ALL_pro_GOid_term)
#this results in a table with protein ID listed in one column and each next column contains a GO ID that was listed with the protein in Uniprot DB.
#convert all factors to  characters
silo39_ALL_pro_GOid_term <- data.frame(lapply(silo39_ALL_pro_GOid_term, as.character), stringsAsFactors = FALSE)
#proteins that don't have GO IDs?

#reshape data so that all GO ID columns are gathered in one column called "GO" 
STACKED_silo39_ALL_pro_GOid_term <- tidyr::gather(silo39_ALL_pro_GOid_term,"col","GO", 2:ncol(silo39_ALL_pro_GOid_term))
#exlude middle column which just contains the string "gene" in each row
STACKED_silo39_ALL_pro_GOid_term <- STACKED_silo39_ALL_pro_GOid_term[,-2]
#remove duplicate rows
STACKED_silo39_ALL_pro_GOid_term <- unique(STACKED_silo39_ALL_pro_GOid_term)
colnames(STACKED_silo39_ALL_pro_GOid_term)[1] <- "protein_ID"
#remove any rows where GO column has NA value. 
STACKED_silo39_ALL_pro_GOid_term <- STACKED_silo39_ALL_pro_GOid_term[which(!is.na(STACKED_silo39_ALL_pro_GOid_term$GO)),]
#this resulting data frame has two columns "gene" and "GO"
#write.csv(STACKED_ALL0.1_pro_GOid_term, "~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/SameDayFCtoTemp/ChiSq_ASCA_clust/intermediate_files/STACKED_ALL0.1_pro_GOid_term.csv", quote = FALSE, row.names = FALSE)

```

### TopGO 
```{r}

###this chunk generates p-values for weight01 fisher test like Liew paper (https://advances.sciencemag.org/content/4/6/eaar8028/tab-figures-data#ref-43)

#background list silo39_ALL_pro_GO
#sig pro list silo39_sig_pro_GO

topgo_bkgd <- silo39_ALL_pro_GO
topgo_bkgd$GO_IDs <- gsub(";",",",topgo_bkgd$GO_IDs)
write.table(topgo_bkgd, "TopGO_bkgd.txt", sep = "\t",quote = FALSE, row.names = FALSE, col.names = FALSE)
geneID2GO_bkgd <- readMappings(file = "TopGO_bkgd.txt")


#to switch between time and temp run line 66 in time and temp functional analysis scripts
topgo_sig <- silo39_sig_pro_GO
topgo_sig$GO_IDs <- gsub(";",",",topgo_sig$GO_IDs)

topgo_sig_1 <- topgo_sig[which(topgo_sig$ClusterID == 1),c("protein_ID", "GO_IDs")]

topgo_sig_2 <- topgo_sig[which(topgo_sig$ClusterID == 2),c("protein_ID", "GO_IDs")]

topgo_sig_3 <- topgo_sig[which(topgo_sig$ClusterID == 3),c("protein_ID", "GO_IDs")]

topgo_sig_4 <- topgo_sig[which(topgo_sig$ClusterID == 4),c("protein_ID", "GO_IDs")]

topgo_sig_5 <- topgo_sig[which(topgo_sig$ClusterID == 5),c("protein_ID", "GO_IDs")]

geneUniverse <- names(geneID2GO_bkgd) 

genesOfInterest_1 <- topgo_sig_1
genesOfInterest_1 <- as.character(genesOfInterest_1$protein_ID)

geneList_1 <- factor(as.integer(geneUniverse %in% genesOfInterest_1))
names(geneList_1) <- geneUniverse

myGOdata_1 <- new("topGOdata", description="time_clade1", ontology="BP", allGenes=geneList_1  ,annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_1 <- runTest(myGOdata_1, algorithm="weight01", statistic="fisher") 

allRes_1 <- GenTable(myGOdata_1, weight = resultFisher_1,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 2

genesOfInterest_2 <- topgo_sig_2
genesOfInterest_2 <- as.character(genesOfInterest_2$protein_ID)

geneList_2 <- factor(as.integer(geneUniverse %in% genesOfInterest_2))
names(geneList_2) <- geneUniverse

myGOdata_2 <- new("topGOdata", description="time_clade2", ontology="BP", allGenes=geneList_2,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_2 <- runTest(myGOdata_2, algorithm="weight01", statistic="fisher") 

allRes_2 <- GenTable(myGOdata_2, weight = resultFisher_2,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 3

genesOfInterest_3 <- topgo_sig_3
genesOfInterest_3 <- as.character(genesOfInterest_3$protein_ID)

geneList_3 <- factor(as.integer(geneUniverse %in% genesOfInterest_3))
names(geneList_3) <- geneUniverse

myGOdata_3 <- new("topGOdata", description="time_clade3", ontology="BP", allGenes=geneList_3,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_3 <- runTest(myGOdata_3, algorithm="weight01", statistic="fisher") 

allRes_3 <- GenTable(myGOdata_3, weight = resultFisher_3,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 4

genesOfInterest_4 <- topgo_sig_4
genesOfInterest_4 <- as.character(genesOfInterest_4$protein_ID)

geneList_4 <- factor(as.integer(geneUniverse %in% genesOfInterest_4))
names(geneList_4) <- geneUniverse

myGOdata_4 <- new("topGOdata", description="time_clade4", ontology="BP", allGenes=geneList_4,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_4 <- runTest(myGOdata_4, algorithm="weight01", statistic="fisher") 

allRes_4 <- GenTable(myGOdata_4, weight = resultFisher_4,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 5

genesOfInterest_5 <- topgo_sig_5
genesOfInterest_5 <- as.character(genesOfInterest_5$protein_ID)

geneList_5 <- factor(as.integer(geneUniverse %in% genesOfInterest_5))
names(geneList_5) <- geneUniverse

myGOdata_5 <- new("topGOdata", description="time_clade5", ontology="BP", allGenes=geneList_5,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_5 <- runTest(myGOdata_5, algorithm="weight01", statistic="fisher") 


allRes_5 <- GenTable(myGOdata_5, weight = resultFisher_5,orderBy = "weight", ranksOf = "classic", topNodes = 50)


sub_allRes1 <- allRes_1[which(allRes_1$Annotated >=5 & allRes_1$weight <= 0.05),]
sub_allRes2 <- allRes_2[which(allRes_2$Annotated >=5 & allRes_2$weight <= 0.05),]
sub_allRes3 <- allRes_3[which(allRes_3$Annotated >=5 & allRes_3$weight <= 0.05),]
sub_allRes4 <- allRes_4[which(allRes_4$Annotated >=5 & allRes_4$weight <= 0.05),]
sub_allRes5 <- allRes_5[which(allRes_5$Annotated >=5 & allRes_5$weight <= 0.05),]


#bind all result subsets into one data frame
sub_allRes <- rbind(sub_allRes1,sub_allRes2,sub_allRes3,sub_allRes4,sub_allRes5)

sub_allRes <- sub_allRes[,1:2]
sub_allRes <- unique(sub_allRes)

sub_allRes <- merge(sub_allRes, sub_allRes1[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[3] <- "clade1"

sub_allRes <- merge(sub_allRes, sub_allRes2[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[4] <- "clade2"

sub_allRes <- merge(sub_allRes, sub_allRes3[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[5] <- "clade3"

sub_allRes <- merge(sub_allRes, sub_allRes4[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[6] <- "clade4"

sub_allRes <- merge(sub_allRes, sub_allRes5[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[7] <- "clade5"

write.csv(sub_allRes, "Time_Clades_SigGO.csv", quote = FALSE, row.names = FALSE)
```

#get GO slims for clade terms
```{r}
#make list of unique GO terms without a protein ID column
GO_clade1 <- sub_allRes[which(!is.na(sub_allRes$clade1)),"GO.ID"]
GO_clade2 <- sub_allRes[which(!is.na(sub_allRes$clade2)),"GO.ID"]
GO_clade3 <- sub_allRes[which(!is.na(sub_allRes$clade3)),"GO.ID"]
GO_clade4 <- sub_allRes[which(!is.na(sub_allRes$clade4)),"GO.ID"]
GO_clade5 <- sub_allRes[which(!is.na(sub_allRes$clade5)),"GO.ID"]



#Use GSEA to generate list of GO Slim BP terms for each clade
#DF will have "term", "GOid", "GOcategory"

#goslims with GSEA
collection_clade1 <- GOCollection(GO_clade1)
collection_clade2 <- GOCollection(GO_clade2)
collection_clade3 <- GOCollection(GO_clade3)
collection_clade4 <- GOCollection(GO_clade4)
collection_clade5 <- GOCollection(GO_clade5)

#I downloaded goslim_generic.obo from http://geneontology.org/docs/go-subset-guide/
#then i moved it to the R library for GSEABase in the extdata folder
fl <- system.file("extdata", "goslim_generic.obo", package="GSEABase")
slim <- getOBOCollection(fl)
slims_clade1 <- data.frame(goSlim(collection_clade1, slim, "BP"))
slims_clade2 <- data.frame(goSlim(collection_clade2, slim, "BP"))
slims_clade3 <- data.frame(goSlim(collection_clade3, slim, "BP"))
slims_clade4 <- data.frame(goSlim(collection_clade4, slim, "BP"))
slims_clade5 <- data.frame(goSlim(collection_clade5, slim, "BP"))

#bind all clade slim terms together
slims <- rbind(slims_clade1, slims_clade2, slims_clade3, slims_clade4, slims_clade5)

slims <- data.frame(unique(slims$Term))
colnames(slims)[1] <- "Term"

slims <- merge(slims, slims_clade1[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[2] <- "clade1"

slims <- merge(slims, slims_clade2[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[3] <- "clade2"

slims <- merge(slims, slims_clade3[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[4] <- "clade3"

slims <- merge(slims, slims_clade4[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[5] <- "clade4"

slims <- merge(slims, slims_clade5[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[6] <- "clade5"

#remove rows without any counts, and biological_process row

slims <- slims[which(apply(slims[,-1],1,sum)!= 0 & slims$Term != "biological_process"),]

write.csv(slims, "Time_Clades_SigGOSlim.csv",quote = F, row.names = F)

slims_m <- slims[,-1]
rownames(slims_m) <- slims$Term
slims_m <- data.matrix(slims_m)
colnames(slims_m) <- gsub("clade","",colnames(slims_m))

#replace truncated terms with full terms
rownames(slims_m) <- gsub("nucleobase-containing compound cata...","nucleobase-containing compound catabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("generation of precursor metabolites...","generation of precursor metabolites and energy", rownames(slims_m))

rownames(slims_m) <- gsub("cytoskeleton-dependent intracellula...","cytoskeleton-dependent intracellular transport", rownames(slims_m))

rownames(slims_m) <- gsub("cellular protein modification proce...","cellular protein modification process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic p", rownames(slims_m))

rownames(slims_m) <- gsub("cellular amino acid metabolic proce...","cellular amino acid metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("anatomical structure formation invo...","anatomical structure formation involved in morphogenesis", rownames(slims_m))




#library(viridis)
#col = rev(viridis(11))
col = c("#FFFFFFFF","#C2DF23FF","#85D54AFF","#51C56AFF","#2BB07FFF","#1E9B8AFF","#25858EFF","#2D708EFF","#38598CFF","#433E85FF","#482173FF","#440154FF")
breaks <- c(-1,0,1,2,3,4,5,6,7,8,9,10,11)

ColSideColors<- c("#000000","#00FF00","#808080","#8000FF","#80FFFF")
pdf(file="GOSlimheatmapTime.pdf", width=8, height=8)
heatmap.2(slims_m,margins = c(5,20),cexRow = 0.8, Colv=NA,revC = TRUE,breaks= breaks,col = col,density.info = "none", trace = "none", scale = "none",ColSideColors = ColSideColors)
dev.off()

breaks <- c(seq(-1,0, length = 2),1,seq(2,11, length = 5))
pdf(file="GOSlimheatmapTime.pdf", width=10, height=12)
heatmap.2(slims_m,margins = c(5,27),cexRow = 1.2, Colv=NA,revC = TRUE,breaks= breaks,col = c("white",viridis(6)),density.info = "none", trace = "none", scale = "none",  hclustfun = function(x) hclust(x, method="average"), distfun = function(x) as.dist(1-cor(t(x), use = "pa")))
dev.off()

```


#######################################
Analysis of Temperature-affected proteins
######################################

**Match GO terms to sig protein lists**

```{r}
#Get GO IDs for silo 3 and 9 sig. proteins
silo39_sig_pro_GO <- merge(silo39_temp_ASCA, uniprot_filtered[,c("protein_ID","GO_IDs")], by ="protein_ID")

#count number of proteins with GO terms
nrow(silo39_sig_pro_GO)

#Since GO IDs are listed all in one column, spread them across multiple columns (1 column /term)
silo39_sig_pro_GOid_term <- data.frame()
for (i in 1:nrow(silo39_sig_pro_GO)){
  silo39_sig_pro_GOid_term_row <- data.frame(t(data.frame(strsplit(as.character(silo39_sig_pro_GO$GO_IDs[i]),'; ', fixed = TRUE))))
  silo39_sig_pro_GOid_term <- rbind.fill(silo39_sig_pro_GOid_term,silo39_sig_pro_GOid_term_row)
}

#add protein IDs back to GO IDs
silo39_sig_pro_GOid_term <- cbind(silo39_sig_pro_GO[,"protein_ID"], silo39_sig_pro_GOid_term)
#this results in a table with protein ID listed in one column and each next column contains a GO ID that was listed with the protein in Uniprot DB.
#convert all factors to  characters
silo39_sig_pro_GOid_term <- data.frame(lapply(silo39_sig_pro_GOid_term, as.character), stringsAsFactors = FALSE)
#proteins that don't have GO IDs?

#reshape data so that all GO ID columns are gathered in one column called "GO" 
STACKED_silo39_sig_pro_GOid_term <- tidyr::gather(silo39_sig_pro_GOid_term,"col","GO", 2:ncol(silo39_sig_pro_GOid_term))
#exlude middle column which just contains the string "gene" in each row
STACKED_silo39_sig_pro_GOid_term <- STACKED_silo39_sig_pro_GOid_term[,-2]
#remove duplicate rows
STACKED_silo39_sig_pro_GOid_term <- unique(STACKED_silo39_sig_pro_GOid_term)
colnames(STACKED_silo39_sig_pro_GOid_term)[1] <- "protein_ID"
#remove any rows where GO column has NA value. 
STACKED_silo39_sig_pro_GOid_term <- STACKED_silo39_sig_pro_GOid_term[which(!is.na(STACKED_silo39_sig_pro_GOid_term$GO)),]
#this resulting data frame has two columns "gene" and "GO"
#write.csv(STACKED_sig0.1_pro_GOid_term, "~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/SameDayFCtoTemp/ChiSq_ASCA_clust/intermediate_files/STACKED_sig0.1_pro_GOid_term.csv", quote = FALSE, row.names = FALSE)
```

### TopGO 
```{r}

###this chunk generates p-values for weight01 fisher test like Liew paper (https://advances.sciencemag.org/content/4/6/eaar8028/tab-figures-data#ref-43)

#background list silo39_ALL_pro_GO
#sig pro list silo39_sig_pro_GO

topgo_bkgd <- silo39_ALL_pro_GO
topgo_bkgd$GO_IDs <- gsub(";",",",topgo_bkgd$GO_IDs)
write.table(topgo_bkgd, "TopGO_bkgd.txt", sep = "\t",quote = FALSE, row.names = FALSE, col.names = FALSE)
geneID2GO_bkgd <- readMappings(file = "TopGO_bkgd.txt")


#to switch between time and temp run line 66 in time and temp functional analysis scripts
topgo_sig <- silo39_sig_pro_GO
topgo_sig$GO_IDs <- gsub(";",",",topgo_sig$GO_IDs)

topgo_sig_1 <- topgo_sig[which(topgo_sig$ClusterID == 1),c("protein_ID", "GO_IDs")]

topgo_sig_2 <- topgo_sig[which(topgo_sig$ClusterID == 2),c("protein_ID", "GO_IDs")]

geneUniverse <- names(geneID2GO_bkgd) 

genesOfInterest_1 <- topgo_sig_1
genesOfInterest_1 <- as.character(genesOfInterest_1$protein_ID)

geneList_1 <- factor(as.integer(geneUniverse %in% genesOfInterest_1))
names(geneList_1) <- geneUniverse

myGOdata_1 <- new("topGOdata", description="temp_clade1", ontology="BP", allGenes=geneList_1  ,annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_1 <- runTest(myGOdata_1, algorithm="weight01", statistic="fisher") 

allRes_1 <- GenTable(myGOdata_1, weight = resultFisher_1,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 2

genesOfInterest_2 <- topgo_sig_2
genesOfInterest_2 <- as.character(genesOfInterest_2$protein_ID)

geneList_2 <- factor(as.integer(geneUniverse %in% genesOfInterest_2))
names(geneList_2) <- geneUniverse

myGOdata_2 <- new("topGOdata", description="temp_clade2", ontology="BP", allGenes=geneList_2,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_2 <- runTest(myGOdata_2, algorithm="weight01", statistic="fisher") 

allRes_2 <- GenTable(myGOdata_2, weight = resultFisher_2,orderBy = "weight", ranksOf = "classic", topNodes = 50)


sub_allRes1 <- allRes_1[which(allRes_1$Annotated >=5 & allRes_1$weight <= 0.05),]
sub_allRes2 <- allRes_2[which(allRes_2$Annotated >=5 & allRes_2$weight <= 0.05),]

#bind all result subsets into one data frame
sub_allRes <- rbind(sub_allRes1,sub_allRes2)

sub_allRes <- sub_allRes[,1:2]
sub_allRes <- unique(sub_allRes)

sub_allRes <- merge(sub_allRes, sub_allRes1[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[3] <- "clade1"

sub_allRes <- merge(sub_allRes, sub_allRes2[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[4] <- "clade2"

write.csv(sub_allRes, "Temp_Clades_SigGO.csv", quote = FALSE, row.names = FALSE)
```

#get GO slims for clade terms
```{r}
#make list of unique GO terms without a protein ID column
GO_clade1 <- sub_allRes[which(!is.na(sub_allRes$clade1)),"GO.ID"]
GO_clade2 <- sub_allRes[which(!is.na(sub_allRes$clade2)),"GO.ID"]

#Use GSEA to generate list of GO Slim BP terms for each clade
#DF will have "term", "GOid", "GOcategory"

#goslims with GSEA
collection_clade1 <- GOCollection(GO_clade1)
collection_clade2 <- GOCollection(GO_clade2)

#I downloaded goslim_generic.obo from http://geneontology.org/docs/go-subset-guide/
#then i moved it to the R library for GSEABase in the extdata folder
#make list of unique GO terms without a protein ID column
#moved it to the R library for GSEABase in the extdata folder
fl <- system.file("extdata", "goslim_generic.obo", package="GSEABase")
slim <- getOBOCollection(fl)
slims_clade1 <- data.frame(goSlim(collection_clade1, slim, "BP"))
slims_clade2 <- data.frame(goSlim(collection_clade2, slim, "BP"))

#bind all clade slim terms together
slims <- rbind(slims_clade1, slims_clade2)

slims <- data.frame(unique(slims$Term))
colnames(slims)[1] <- "Term"

slims <- merge(slims, slims_clade1[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[2] <- "clade1"

slims <- merge(slims, slims_clade2[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[3] <- "clade2"

#remove rows without any counts, and biological_process row

slims <- slims[which(apply(slims[,-1],1,sum)!= 0 & slims$Term != "biological_process"),]

write.csv(slims, "Temp_Clades_SigGOSlim.csv",quote = F, row.names = F)

slims_m <- slims[,-1]
rownames(slims_m) <- slims$Term
slims_m <- data.matrix(slims_m)
colnames(slims_m) <- gsub("clade","",colnames(slims_m))

#replace truncated terms with full terms
rownames(slims_m) <- gsub("nucleobase-containing compound cata...","nucleobase-containing compound catabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("generation of precursor metabolites...","generation of precursor metabolites and energy", rownames(slims_m))

rownames(slims_m) <- gsub("cytoskeleton-dependent intracellula...","cytoskeleton-dependent intracellular transport", rownames(slims_m))

rownames(slims_m) <- gsub("cellular protein modification proce...","cellular protein modification process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic p", rownames(slims_m))

rownames(slims_m) <- gsub("cellular amino acid metabolic proce...","cellular amino acid metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("anatomical structure formation invo...","anatomical structure formation involved in morphogenesis", rownames(slims_m))


heatmap.2(slims_m,margins = c(5,27),cexRow = 1.2, Colv=NA,revC = TRUE,col = c("white",viridis(6)),density.info = "none", trace = "none", scale = "none")


breaks <- c(0,0.5,1,2,3,4,5)
pdf(file="GOSlimheatmapTemp.pdf", width=8, height=10)
heatmap.2(slims_m,margins = c(5,25),cexRow = 1.2, Colv=NA,revC = TRUE,breaks= breaks,col = c("white",viridis(5)),density.info = "none", trace = "none", scale = "none")
dev.off()

```

#######################################
Analysis of Day21 and day27 affected proteins
######################################

**Match GO terms to sig protein lists**

```{r}
silo39_global_PCA <- read.csv("../ASCA/PCA_d21d27.csv", stringsAsFactors = FALSE)
colnames(silo39_temp_ASCA)[1] <- "protein_ID"

#select only black and khaki clades
silo39_global_PCA <- silo39_global_PCA[which(silo39_global_PCA$ClusterID == 1 | silo39_global_PCA$ClusterID==3),]


#Get GO IDs for silo 3 and 9 sig. proteins
silo39_sig_pro_GO <- merge(silo39_global_PCA, uniprot_filtered[,c("protein_ID","GO_IDs")], by ="protein_ID")

#count number of proteins with GO terms
nrow(silo39_sig_pro_GO)

#Since GO IDs are listed all in one column, spread them across multiple columns (1 column /term)
silo39_sig_pro_GOid_term <- data.frame()
for (i in 1:nrow(silo39_sig_pro_GO)){
  silo39_sig_pro_GOid_term_row <- data.frame(t(data.frame(strsplit(as.character(silo39_sig_pro_GO$GO_IDs[i]),'; ', fixed = TRUE))))
  silo39_sig_pro_GOid_term <- rbind.fill(silo39_sig_pro_GOid_term,silo39_sig_pro_GOid_term_row)
}

#add protein IDs back to GO IDs
silo39_sig_pro_GOid_term <- cbind(silo39_sig_pro_GO[,"protein_ID"], silo39_sig_pro_GOid_term)
#this results in a table with protein ID listed in one column and each next column contains a GO ID that was listed with the protein in Uniprot DB.
#convert all factors to  characters
silo39_sig_pro_GOid_term <- data.frame(lapply(silo39_sig_pro_GOid_term, as.character), stringsAsFactors = FALSE)
#proteins that don't have GO IDs?

#reshape data so that all GO ID columns are gathered in one column called "GO" 
STACKED_silo39_sig_pro_GOid_term <- tidyr::gather(silo39_sig_pro_GOid_term,"col","GO", 2:ncol(silo39_sig_pro_GOid_term))
#exlude middle column which just contains the string "gene" in each row
STACKED_silo39_sig_pro_GOid_term <- STACKED_silo39_sig_pro_GOid_term[,-2]
#remove duplicate rows
STACKED_silo39_sig_pro_GOid_term <- unique(STACKED_silo39_sig_pro_GOid_term)
colnames(STACKED_silo39_sig_pro_GOid_term)[1] <- "protein_ID"
#remove any rows where GO column has NA value. 
STACKED_silo39_sig_pro_GOid_term <- STACKED_silo39_sig_pro_GOid_term[which(!is.na(STACKED_silo39_sig_pro_GOid_term$GO)),]
#this resulting data frame has two columns "gene" and "GO"
#write.csv(STACKED_sig0.1_pro_GOid_term, "~/Documents/GitHub/OysterSeedProject/analysis/UniprotAnnotations_NetworkAnalysis/SameDayFCtoTemp/ChiSq_ASCA_clust/intermediate_files/STACKED_sig0.1_pro_GOid_term.csv", quote = FALSE, row.names = FALSE)
```

### TopGO 
```{r}

###this chunk generates p-values for weight01 fisher test like Liew paper (https://advances.sciencemag.org/content/4/6/eaar8028/tab-figures-data#ref-43)

#background list silo39_ALL_pro_GO
#sig pro list silo39_sig_pro_GO

topgo_bkgd <- silo39_ALL_pro_GO
topgo_bkgd$GO_IDs <- gsub(";",",",topgo_bkgd$GO_IDs)
write.table(topgo_bkgd, "TopGO_bkgd.txt", sep = "\t",quote = FALSE, row.names = FALSE, col.names = FALSE)
geneID2GO_bkgd <- readMappings(file = "TopGO_bkgd.txt")


#to switch between time and temp run line 66 in time and temp functional analysis scripts
topgo_sig <- silo39_sig_pro_GO
topgo_sig$GO_IDs <- gsub(";",",",topgo_sig$GO_IDs)

topgo_sig_1 <- topgo_sig[which(topgo_sig$ClusterID == 1),c("protein_ID", "GO_IDs")]

topgo_sig_2 <- topgo_sig[which(topgo_sig$ClusterID == 3),c("protein_ID", "GO_IDs")]

geneUniverse <- names(geneID2GO_bkgd) 

genesOfInterest_1 <- topgo_sig_1
genesOfInterest_1 <- as.character(genesOfInterest_1$protein_ID)

geneList_1 <- factor(as.integer(geneUniverse %in% genesOfInterest_1))
names(geneList_1) <- geneUniverse

myGOdata_1 <- new("topGOdata", description="glob_clade1", ontology="BP", allGenes=geneList_1  ,annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_1 <- runTest(myGOdata_1, algorithm="weight01", statistic="fisher") 

allRes_1 <- GenTable(myGOdata_1, weight = resultFisher_1,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 2

genesOfInterest_2 <- topgo_sig_2
genesOfInterest_2 <- as.character(genesOfInterest_2$protein_ID)

geneList_2 <- factor(as.integer(geneUniverse %in% genesOfInterest_2))
names(geneList_2) <- geneUniverse

myGOdata_2 <- new("topGOdata", description="glob_clade2", ontology="BP", allGenes=geneList_2,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_2 <- runTest(myGOdata_2, algorithm="weight01", statistic="fisher") 

allRes_2 <- GenTable(myGOdata_2, weight = resultFisher_2,orderBy = "weight", ranksOf = "classic", topNodes = 50)


sub_allRes1 <- allRes_1[which(allRes_1$Annotated >=5 & allRes_1$weight <= 0.05),]
sub_allRes2 <- allRes_2[which(allRes_2$Annotated >=5 & allRes_2$weight <= 0.05),]

#bind all result subsets into one data frame
sub_allRes <- rbind(sub_allRes1,sub_allRes2)

sub_allRes <- sub_allRes[,1:2]
sub_allRes <- unique(sub_allRes)

sub_allRes <- merge(sub_allRes, sub_allRes1[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[3] <- "clade1"

sub_allRes <- merge(sub_allRes, sub_allRes2[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[4] <- "clade2"

```

#get GO slims for clade terms
```{r}

GO_clade1 <- sub_allRes[which(!is.na(sub_allRes$clade1)),"GO.ID"]
GO_clade2 <- sub_allRes[which(!is.na(sub_allRes$clade2)),"GO.ID"]

#Use GSEA to generate list of GO Slim BP terms for each clade
#DF will have "term", "GOid", "GOcategory"

#goslims with GSEA
collection_clade1 <- GOCollection(GO_clade1)
collection_clade2 <- GOCollection(GO_clade2)

#make list of unique GO terms without a protein ID column
#moved it to the R library for GSEABase in the extdata folder
fl <- system.file("extdata", "goslim_generic.obo", package="GSEABase")
slim <- getOBOCollection(fl)
slims_clade1 <- data.frame(goSlim(collection_clade1, slim, "BP"))
slims_clade2 <- data.frame(goSlim(collection_clade2, slim, "BP"))

#bind all clade slim terms together
slims <- rbind(slims_clade1, slims_clade2)

slims <- data.frame(unique(slims$Term))
colnames(slims)[1] <- "Term"

slims <- merge(slims, slims_clade1[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[2] <- "clade1"

slims <- merge(slims, slims_clade2[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[3] <- "clade2"

#remove rows without any counts, and biological_process row

slims <- slims[which(apply(slims[,-1],1,sum)!= 0 & slims$Term != "biological_process"),]


slims_m <- slims[,-1]
rownames(slims_m) <- slims$Term
slims_m <- data.matrix(slims_m)
colnames(slims_m) <- gsub("clade","",colnames(slims_m))

#replace truncated terms with full terms
rownames(slims_m) <- gsub("nucleobase-containing compound cata...","nucleobase-containing compound catabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("generation of precursor metabolites...","generation of precursor metabolites and energy", rownames(slims_m))

rownames(slims_m) <- gsub("cytoskeleton-dependent intracellula...","cytoskeleton-dependent intracellular transport", rownames(slims_m))

rownames(slims_m) <- gsub("cellular protein modification proce...","cellular protein modification process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic p", rownames(slims_m))

rownames(slims_m) <- gsub("cellular amino acid metabolic proce...","cellular amino acid metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("anatomical structure formation invo...","anatomical structure formation involved in morphogenesis", rownames(slims_m))

breaks <- c(seq(-1,0, length = 2),1,seq(2,8, length = 5))
heatmap.2(slims_m,margins = c(5,27),cexRow = 1.2, Colv=NA,revC = TRUE,density.info = "none", trace = "none",col = c("white", viridis(8)), scale = "none")
