---
title: "PCA_ASCA_Functional_Analysis"
author: "Shelly Trigg"
date: "9/13/2019"
output: rmarkdown::github_document
editor_options: 
  chunk_output_type: console
---

**load libraries**
```{r, echo = FALSE}
library(dplyr)
library(tidyr)
library(MetStaT)
library(ggplot2)
library(heatmap3)
library(colorRamps)
library(plotrix)
library(plyr)
library(GSEABase)
library(reshape2)
library(gplots)
library(viridis)
library(topGO)
```

**load NSAF data**
```{r}
#Read in NSAF data 

data <- read.csv("../../Data/Average_adjNSAF_nozerovals.csv", stringsAsFactors = FALSE)

```


## PCA Analysis

**Run PCA**
```{r}
#log transform NSAF values
data_log <-  log(data[,-c(1:2)],2)

#create PCA object
set.seed(101)
pca_log <- prcomp(data_log, center = T, scale = T)
summary_pca_log <- summary(pca_log)
summary_pca_log
capture.output(summary(pca_log), file = "PCA_model_summary.txt")

#add meta data to PCA data
pca_log_meta <- cbind(data$dpf, data$temp, data.frame(paste(data$dpf,data$temp, sep = "_")),pca_log$x)
#rename columns
colnames(pca_log_meta)[1:3] <- c("dpf","temp","SampleName")
```

Plot PC1 x PC2 of PCA on averaged NSAF values with days ordered from earliest to latest
```{r}
jpeg("../../Figures/Figure2a.jpg", height = 5, width = 7, units = "in", res = 300 )
ggplot(pca_log_meta, aes(PC1, PC2,label = dpf, color = as.factor(temp))) + geom_text(size = 5, fontface = "bold") + scale_colour_manual(values = c("brown4","cyan3", "magenta3")) + theme_bw() + ylab(paste("PC2"," (",formatC(summary_pca_log$importance[5] * 100,digits=2,format="f"),"%)", sep = "")) + xlab(paste("PC1"," (",formatC(summary_pca_log$importance[2] * 100,digits=2,format="f"),"%)", sep = ""))
dev.off()

```

Apply threshold to proteins based on PC1 and PC2 loadings
```{r}
#extract loadings for PC1 and PC2
PC1_great <- data.frame(pca_log$rotation)
PC1_great <- PC1_great[which(PC1_great$PC1 > 0),1:2]

#export plot
jpeg("../../SupplementalFiles/FigureS2a.jpg", height = 5, width = 5, units = "in", res = 300 )
plot(head(PC1_great[order(PC1_great, decreasing = TRUE),1],100), xlab = "protein", ylab = "PC1 Loadings")
abline(h=0.0236, col = "red")
dev.off()


PC1_great_prots <- rownames(PC1_great[which(PC1_great$PC1 > 0.0236),])

PC1_less <- data.frame(pca_log$rotation)
PC1_less <- PC1_less[which(PC1_less$PC1 < 0),1:2]

jpeg("../../SupplementalFiles/FigureS2b.jpg", height = 5, width = 5, units = "in", res = 300 )
plot(tail(PC1_less[order(PC1_less$PC1, decreasing = TRUE),1],100), xlab = "protein", ylab = "PC1 Loadings")
abline(h=-0.02355, col = "red")
dev.off()

PC1_less_prots <- rownames(PC1_less[which(PC1_less$PC1 < -0.02355),])

#now PC2
PC2_great <- data.frame(pca_log$rotation)
PC2_great <- PC2_great[which(PC2_great$PC2 > 0),1:2]

jpeg("../../SupplementalFiles/FigureS2c.jpg", height = 5, width = 5, units = "in", res = 300 )
plot(head(PC2_great[order(PC2_great$PC2, decreasing = TRUE),2],100), xlab = "protein", ylab = "PC2 Loadings")
abline(h=0.0245, col = "red")
dev.off()

PC2_great_prots <- rownames(PC2_great[which(PC2_great$PC2 > 0.0245),])

PC2_less <- data.frame(pca_log$rotation)
PC2_less <- PC2_less[which(PC2_less$PC2 < 0),1:2]

jpeg("../../SupplementalFiles/FigureS2d.jpg", height = 5, width = 5, units = "in", res = 300 )
plot(tail(PC2_less[order(PC2_less$PC2, decreasing = TRUE),2],100), xlab = "protein", ylab = "PC2 Loadings")
abline(h=-0.0262, col = "red")
dev.off()

PC2_less_prots <- rownames(PC2_less[which(PC2_less$PC2 < -0.0262),])


global_PC12 <- data.frame(c(PC1_great_prots,PC1_less_prots, PC2_great_prots, PC2_less_prots))

colnames(global_PC12)[1] <- "protein" 

```

Format table with all proteins and their loadings values for PC1 and PC2
```{r}
PC_loadings <- data.frame(cbind(rownames(pca_log$rotation), pca_log$rotation[,1],pca_log$rotation[,2]))

#add column names
colnames(PC_loadings) <- c("protein_ID", "PC1_loadings", "PC2_loadings")

#remove rownames
rownames(PC_loadings) <- NULL

PC_loadings$PC_threshold_pass <- "No"

for(i in 1:nrow(PC_loadings)){
  if(PC_loadings[i,"protein_ID"] %in% global_PC12$protein){
    PC_loadings$PC_threshold_pass[i] <- "Yes"
    }
  else(PC_loadings$PC_threshold_pass[i] <- "No")
}

```

Plot heatmap for global PCA (Fig2a) PC1 and PC2 biggest contributors
```{r}
#rename column 1
colnames(global_PC12)[1] <- "protein"
#replace pipe in protein names with decimal
global_PC12$protein <- gsub("\\|", "\\.", global_PC12$protein)
#subset NSAF data for global PCA PC1 and PC2 biggest contributors
global_cut_data <- data[,c(1:2,which(colnames(data) %in% global_PC12$protein))]

#order the data by day then temp
global_cut_data <- global_cut_data[order(global_cut_data$dpf,global_cut_data$temp),]
#add temp and day info to column names
rownames(global_cut_data) <- paste(global_cut_data$dpf, global_cut_data$temp, sep="_")
#remove day and temp columns
global_cut_data <- global_cut_data[,-c(1:2)]

#save a color object for temperature labels
ColSideColors<-cbind(Temp=c("brown",rep(c("cyan3","magenta2"),6)))

#plot the heat map

heatmap3(as.matrix(t(global_cut_data)), Colv = NA, cexRow = 0.5,labCol = c(19,21,21,23,23,25,25,27,27,29,29,31,31), ColSideColors = ColSideColors, ColAxisColors = 1, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), method = "complete")

#what are the proteins that distinguish day 21 and day 27?
hm <- as.dist(1-cor(as.matrix(global_cut_data), use="pa"))
hm2 <- hclust(hm, method = 'complete')
plot(hm2, cex=0.3)
abline(h=1, col = "red")

#Identify clades in heatmap
# define some clusters

#make a list of proteins with cluster IDs based on where the tree is cut
mycl <- cutree(hm2, h=1)

#assign colors to cluster IDs
clusterCols <- rev(RColorBrewer::brewer.pal(n = length(unique(mycl)), "Set3"))

myClusterSideBar <- clusterCols[mycl]

#get color name
clusterColor <- lapply(myClusterSideBar, color.id)
clusterColor <- lapply(clusterColor, `[[`,1)
clusterColor <- data.frame(unlist(clusterColor), stringsAsFactors = FALSE)

foo <- cbind(data.frame(mycl), clusterColor)
#add protein column to data frame for merging
foo$protein_ID <- rownames(foo)

#add protein column back to NSAF data frame for merging
global_cut_data_proteinID <- data.frame(t(global_cut_data))
global_cut_data_proteinID$protein_ID <- rownames(global_cut_data_proteinID)

clade.prot <- merge(foo,global_cut_data_proteinID,  by = "protein_ID")

colnames(clade.prot)[3] <- "ClusterColor"

#simplify clade names
clade.prot$SimpleClusterColor <- clade.prot$ClusterColor
clade.prot$SimpleClusterColor <- gsub("lightsteelblue","pale purple", clade.prot$SimpleClusterColor)
clade.prot$SimpleClusterColor <- gsub("wheat1","pale yellow", clade.prot$SimpleClusterColor)


#plot heatmap with colored row side bars
heatmap3(as.matrix(t(global_cut_data)), Colv = NA, cexRow = 0.5,labCol = c(19,21,21,23,23,25,25,27,27,29,29,31,31), ColSideColors = ColSideColors,RowSideColors=myClusterSideBar, ColAxisColors = 1,RowAxisColors=1,distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), method = "complete")

#teal clade does not show much difference between temperatures
#exclude "teal" cluster that does not show much change between temperatures 

jpeg(file="../../Figures/Figure2b.jpg.jpg", width=7, height=5, units = "in", res = 300)
heatmap3(as.matrix(t(global_cut_data)[which(!(rownames(t(global_cut_data)) %in% clade.prot[which(clade.prot$ClusterColor =="paleturquoise3"),"protein_ID"])),]), Colv = NA, labRow = NA,labCol = c(19,21,21,23,23,25,25,27,27,29,29,31,31), ColSideColors = ColSideColors,RowSideColors=myClusterSideBar[myClusterSideBar != "#8DD3C7"], ColAxisColors = 1,RowAxisColors=1,distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), method = "complete")
dev.off()

```

Merge cluster color and id with PC loadings data
```{r}
PC_loadings <- merge(PC_loadings, clade.prot[,c("protein_ID", "ClusterColor", "SimpleClusterColor")], by = c("protein_ID"), all.x = TRUE)

```

## ASCA ANALYSIS

make a dataframe with just CHOYP proteins and exclude day 0 for data going into ASCA
```{r subset_data_for_CHOYP_only_rm0, echo = FALSE}
choyp_data <- data[which(data$dpf !=19),c(1:2,grep("CHOYP", colnames(data)))]
```

**Perform ASCA on log2 data**
```{r}
#create matrix to pass to ASCA command, excluding the silo and time info
ASCA_X <- as.matrix(log(choyp_data[,-c(1:2)],2))
#create matrix to pass to ASCA command with only the silo and time info
ASCA_F <- as.matrix(choyp_data[,c(1:2)])
#perform ASCA
set.seed(101)
ASCA <- ASCA.Calculate(ASCA_X, ASCA_F, equation.elements = "1,2,12", scaling = FALSE)
```

Here is a summary of the ASCA results (e.g. variance explained by different factors; factor 1= time (dpf), factor 2 = temperature, interaction = interaction of time and temperature)
```{r}
#print the ASCA summary
ASCA.GetSummary(ASCA)
#write results out to a file
capture.output(ASCA.GetSummary(ASCA), file = "ASCA_model_summary.txt")
```

Perform a permutation test on ASCA results to see if experimental effects are significant beyond randomly permuted data
```{r}
set.seed(101)
perm.test <- ASCA.DoPermutationTest(ASCA,perm = 10000)
capture.output(perm.test, file = "ASCA_permTest_summary.txt")
```


### Plot PCAs from ASCA  

**This first plot is the time (dpf) effect PCA**
```{r avgNSAF_PCA_timeEffect_plot}
#plot PCA for factor 1, which is time in this case
time_PC12 <- data.frame(ASCA$`1`$svd$t[,c(1,2)])
time_PC12 <- cbind(data.frame(sort(rep(ASCA$`1`$level.combinations$row.patterns,2))),data.frame(rep(c("23","29"),6)) , time_PC12)
colnames(time_PC12)<- c("dpf","temp","PC1","PC2")
time_PC12$dpf <- factor(time_PC12$dpf)
time_PC12$PC1_j <- jitter(time_PC12$PC1)
time_PC12$PC2_j <- jitter(time_PC12$PC2)

jpeg(file="../../Figures/Figure3a.jpg", width=7, height=5, units = "in", res = 300)
ggplot(time_PC12, aes(PC1_j, PC2_j, label = dpf, color = temp)) + geom_text(size = 6, fontface = "bold") + scale_color_manual(values = c("cyan3", "magenta3"))+ theme(panel.grid.major =element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(),axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA, size=1), text = element_text(size=15)) + xlab(paste("PC1"," (",formatC(ASCA$`1`$svd$var.explained[1] * 100,digits=2,format="f"),"%)", sep = "")) + ylab(paste("PC2"," (",formatC(ASCA$`1`$svd$var.explained[2] * 100,digits=2,format="f"),"%)", sep = ""))
dev.off()
```

**This next plot is the temperature effect PCA**
```{r avgNSAF_PCA_tempEffect_plot}
#plot PCA for factor 2, which is temperature in this case
temp_PC12 <- data.frame(ASCA$`2`$svd$t[,c(1,2)])
temp_PC12 <- cbind(data.frame(sort(rep(ASCA$`1`$level.combinations$row.patterns,2))),data.frame(rep(c("23","29"),6)) , temp_PC12)
colnames(temp_PC12)<- c("dpf","temp","PC1","PC2")
temp_PC12$dpf <- as.factor(temp_PC12$dpf)


temp_PC12$PC1_j <- jitter(temp_PC12$PC1)
temp_PC12$PC2_j <- jitter(temp_PC12$PC2)

jpeg(file="../../Figures/Figure5a.jpg", width=7, height=5, units = "in", res = 300)
ggplot(temp_PC12, aes(PC1_j, PC2_j, label = dpf, color = temp)) + geom_text(size = 3, fontface = "bold") + scale_color_manual(values = c("cyan3", "magenta3")) + theme(panel.grid.major =element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(),axis.line = element_line(colour = "black"),panel.border = element_rect(colour = "black", fill=NA, size=1), text = element_text(size=10)) + xlab(paste("PC1"," (",formatC(ASCA$`2`$svd$var.explained[1] * 100,digits=2,format="f"),"%)", sep = "")) + ylab(paste("PC2"," (",formatC(ASCA$`2`$svd$var.explained[2] * 100,digits=2,format="f"),"%)", sep = ""))
dev.off()
```


### Analysis of proteins affected by temperature

Because the temperature effect PCA show the most separation between 23C and 29C in PC2, we will look at those loadings.

**PC1 loadings for temperature effect**
```{r avgNSAF_PCA_tempEffect_PC2loadings, echo = FALSE}
#extract protein names from ASCA data; these will be combined with loadings values; the order is maintained
protnames <- data.frame(colnames(ASCA$data))
#combine protein names with ASCA loadings for PC2 for temperature, since this component showed the greatest separation between temperatures
d <- cbind(protnames, ASCA$`2`$svd$v[,1])
#rename the columns
colnames(d) <- c("protein", "PC1loadings")
#make a dataframe of proteins with PC loadings greater than zero for the loadings plot
d_great <- d[which(d$PC1loadings > 0),]
#make a dataframe of proteins with PC loadings less than zero for the loadings plot
d_less <- d[which(d$PC1loadings < 0),]


#plot PC1 loadings
jpeg(file="../../SupplementalFiles/FigureS3e.jpg", width=10, height=8, units = "in", res = 300)
plot(d_great[order(d_great$PC1loadings, decreasing = TRUE),2], xlab = "protein", ylab = "PC1 Loadings")
abline(h=0.03, col = "red")
dev.off()

jpeg(file="../../SupplementalFiles/FigureS3f.jpg", width=10, height=8, units = "in", res = 300)
plot(d_less[order(d_less$PC1loadings, decreasing = TRUE),2],xlab = "protein", ylab = "PC1 Loadings")
abline(h=-0.025, col = "red")
dev.off()

```

To pull out proteins affected by temperature based on their influence in seperating treatment groups on PC1 of the temperature PCA, I picked an absolute value loadings threshold of 0.025. This means any protein that had a loadings value > 0.03 or < -0.025 was selected.
```{r, echo = FALSE}
#make list of proteins with temperature PC1 loadings values >= 0.025
cutd <- d[which(d$PC1loadings >= 0.03 | d$PC1loadings <= -0.025),]
#make a list of cutoff proteins with normalized abundance
cut_data <-data[,which(colnames(data) %in% cutd$protein)]
rownames(cut_data) <- paste(data$dpf, data$temp, sep="_")
```


Number of proteins affected by temperature at loadings value > 0.03 or < -0.025
```{r, echo = FALSE}
nrow(cutd)
```



**Format matrix for heatmap of proteins affected by temperature based on temperature effect PC1 loadings value cutoff**
```{r}
#reorder the cut data so that samples are grouped by temperature
cut_data_ord <- data[,c(1:3,which(colnames(data) %in% cutd$protein))]
cut_data_ord <- cut_data_ord[order(cut_data_ord$temp,cut_data_ord$dpf),]
rownames(cut_data_ord) <- paste(cut_data_ord$dpf, cut_data_ord$temp, sep="_")
cut_data_ord <- cut_data_ord[,-c(1:3)]
#transposed to eventually get row names as columns to match entry names to. 
#t.data.frame returns a matrix so I had convert back to data frame for all downstream rearranging because data frames are much easier to work with
cut_data_ord_t <- data.frame(t.data.frame(cut_data_ord))
#make row names which contain the protein IDs a column
cut_data_ord_t <- cbind(rownames(cut_data_ord_t), cut_data_ord_t)
#add the column name
colnames(cut_data_ord_t)[1] <- "Protein.ID"
#convert the Protein ID column to character
cut_data_ord_t[,1] <- as.character(cut_data_ord_t[,1])
#remove the protein ID column since the row names contain that info
cut_data_ord_t_m <- cut_data_ord_t[,-1]
#this step removes the x from the column names while preserving the order of colnames
colnames(cut_data_ord_t_m) <- rownames(cut_data_ord)

```

**Perform clustering and determine cluster number for heatmap**
```{r}
hm <- as.dist(1-cor(t.data.frame(cut_data_ord_t_m), use="pa"))
hm2 <- hclust(hm, method = 'complete')
plot(hm2, cex=0.3)
abline(h=1.85, col = "red")

#Identify clades in heatmap
# define some clusters

#make a list of proteins with cluster IDs based on where the tree is cut
#1.8 gives 2 clusters
mycl <- cutree(hm2, h=1.85)

#assign colors to cluster IDs
clusterCols <- head(RColorBrewer::brewer.pal(length(unique(mycl)),"Dark2"), length(unique(mycl)))
myClusterSideBar <- clusterCols[mycl]


#make data frames to store heatmap color values
#columns are time/temp; color 23C blue, 29C red, each dpf a different shade of gray
ColSideColors<- cbind(temp = c("brown",rep("cyan3",6),rep("magenta2",6)))


#plot heat map 
jpeg(file="../../Figures/Figure5b.jpg", width=8, height=8, units = "in", res = 300)
heatmap3(as.matrix(cut_data_ord_t_m),cexCol = 1.5, RowSideColors=clusterCols[mycl],ColSideColors = ColSideColors,RowAxisColors=1,ColAxisColors=1,hclustfun=hclust, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), method = "complete", scale = "row", Colv=NA,labRow = NA, RowSideLabs = NA, labCol = c(19,21,23,25,27,29,31,21,23,25,27,29,31))
dev.off()

#get color name
clusterColor <- lapply(myClusterSideBar, color.id)
clusterColor <- lapply(clusterColor, `[[`,1)
clusterColor <- data.frame(unlist(clusterColor), stringsAsFactors = FALSE)

foo <- cbind(data.frame(mycl), clusterColor)
#add protein column to data frame for merging
foo$protein_ID <- rownames(foo)

#add protein column back to NSAF data frame for merging
cut_data_ord_t_m_proteinID <- cut_data_ord_t_m
cut_data_ord_t_m_proteinID$protein_ID <- rownames(cut_data_ord_t_m)

clade.prot <- merge(foo, cut_data_ord_t_m_proteinID, by = "protein_ID")

colnames(clade.prot)[3] <- "ClusterColor"


#simplify clade names
clade.prot$SimpleClusterColor <- clade.prot$ClusterColor
clade.prot$SimpleClusterColor <- gsub("darkorange3","orange", clade.prot$SimpleClusterColor)
clade.prot$SimpleClusterColor <- gsub("cyan4","dark teal", clade.prot$SimpleClusterColor)


#remove day 0 from clade prot
clade.prot_noD0 <- clade.prot[,-4]

#normalize within terms (autoscale by row)
#calculate row means and SD
clade.prot_noD0$mean <- apply(clade.prot_noD0[,4:15], 1, mean)
clade.prot_noD0$sd <- apply(clade.prot_noD0[,4:15],1, sd)

#sweep second argument is margin which corresponds to 1 (rows) or 2 (columns); this is for autoscaling by row to make colors relative to one another like the heatmap does
#subtract the row mean from each value
clade.prot_no_D0_norm <- sweep(clade.prot_noD0[,4:15],1,clade.prot_noD0$mean)
#divide each value by the row standard deviation
clade.prot_no_D0_norm <- sweep(clade.prot_no_D0_norm[,1:12],1,clade.prot_noD0$sd, FUN = "/")

clade.prot_no_D0_norm <- cbind(clade.prot_noD0[,c("protein_ID", "SimpleClusterColor","ClusterColor")], clade.prot_no_D0_norm)

STACKED_NSAF <- tidyr::gather(clade.prot_no_D0_norm, dpftemp, NSAF, 4:15)
STACKED_NSAF$dpf <- gsub("X|_.*","",STACKED_NSAF$dpftemp)
STACKED_NSAF$temp <- as.factor(gsub(".*_","",STACKED_NSAF$dpftemp))
str(STACKED_NSAF)
STACKED_NSAF$dpf <- as.integer(STACKED_NSAF$dpf)

#make mean summary table for plotting summary lines of protein abundances/clade
STACKED_NSAF_mean <- STACKED_NSAF %>%
    dplyr::group_by(dpf, temp, SimpleClusterColor,ClusterColor) %>%
    dplyr::summarise(n = n(), mean=mean(NSAF), median=median(NSAF), sd=sd(NSAF)) %>%
    dplyr::mutate(sem = sd / sqrt(n - 1),
      CI_lower = mean + qt((1-0.95)/2, n - 1) * sem,
      CI_upper = mean - qt((1-0.95)/2, n - 1) * sem)

jpeg("../../Figures/Figure5c.jpg", width=10, height=5, units = "in", res = 300)
ggplot(STACKED_NSAF_mean, aes(x=dpf, y=mean, color = temp)) + geom_line(aes(x=dpf, y=mean, color=temp)) + ylab("mean autoscaled NSAF") + geom_ribbon(aes(ymin=CI_lower,ymax=CI_upper,fill=temp),color="grey70",alpha=0.4)  + facet_wrap(~SimpleClusterColor) + scale_color_manual(values=c("cyan3", "magenta2")) + scale_fill_manual(values=c("cyan3", "magenta2")) + scale_x_continuous(breaks = c(21,23,25,27,29,31), labels = c(21,23,25,27,29,31)) + theme_bw() + theme(text = element_text(size=30),strip.text = element_text(face = "bold",size = 30),strip.background = element_rect(fill="white"))
dev.off()



#create simple line plots for showing main trends over time for temp clades
jpeg(file="../../Figures/Figure6_colLabs.jpg", width=5, height=2, units = "in", res = 300)
ggplot(STACKED_NSAF_mean, aes(x=dpf, y=mean, color = temp)) + geom_line(aes(x=dpf, y=mean, color=temp, size = 1)) + facet_wrap(~SimpleClusterColor, ncol = 3) + scale_color_manual(values=c("cyan3", "magenta2")) + scale_x_continuous(breaks = c(21,23,25,27,29,31), labels = c(21,23,25,27,29,31)) + theme_bw() + theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),strip.text = element_text(face = "bold",size = 32),strip.background = element_rect(fill="white"))
dev.off()

```

Get ASCA temperature PC1 loadings and cluster info into a table
Format table with all proteins and their loadings values for PC1 and PC2
```{r}
temp_ASCA_PC1_loadings <- d

#add column names
colnames(temp_ASCA_PC1_loadings) <- c("protein_ID", "ASCA_PC1_loadings")

temp_ASCA_PC1_loadings$ASCA_PC_threshold_pass <- "No"

for(i in 1:nrow(temp_ASCA_PC1_loadings)){
  if(temp_ASCA_PC1_loadings[i,"protein_ID"] %in% cutd$protein){
    temp_ASCA_PC1_loadings$ASCA_PC_threshold_pass[i] <- "Yes"
    }
  else(temp_ASCA_PC1_loadings$ASCA_PC_threshold_pass[i] <- "No")
}

#Merge cluster color and id with PC loadings data
temp_ASCA_PC1_loadings <- merge(temp_ASCA_PC1_loadings, clade.prot[,c("protein_ID", "ClusterColor", "SimpleClusterColor")], by = "protein_ID", all.x = TRUE)


```

## PART 2
### Analysis of proteins affected by time 

**PC1 loadings for the time effect PCA**
```{r avgNSAF_PCA_timeEffect_PC1loadings, echo = FALSE}
#make dataframes of factor 1 (time) PC1 loadings
d1 <- cbind(protnames, ASCA$`1`$svd$v[,1])
colnames(d1) <- c("protein", "PC1loadings")
d1_great <- d1[which(d1$PC1loadings > 0),]
d1_less <- d1[which(d1$PC1loadings < 0),]

jpeg(file="../../SupplementalFiles/FigureS3a.jpg", width=10, height=8, units = "in", res = 300)
plot(d1_great[order(d1_great$PC1loadings, decreasing = TRUE),2], ylab = "PC1 Loadings")
abline(h=0.035, col = "red")
dev.off()

jpeg(file="../../SupplementalFiles/FigureS3b.jpg", width=10, height=8, units = "in", res = 300)
plot(d1_less[order(d1_less$PC1loadings, decreasing = TRUE),2],ylab = "PC1 Loadings")
abline(h=-0.035, col = "red")
dev.off()
```

**PC2 loadings for the time effect PCA**
```{r avgNSAF_PCA_timeEffect_PC2loadings, echo = FALSE}
d2 <- cbind(protnames, ASCA$`1`$svd$v[,2])
colnames(d2) <- c("protein", "PC2loadings")
d2_great <- d2[which(d2$PC2loadings > 0),]
d2_less <- d2[which(d2$PC2loadings < 0),]

jpeg(file="../../SupplementalFiles/FigureS3c.jpg", width=10, height=8, units = "in", res = 300)
plot(d2_great[order(d2_great$PC2loadings, decreasing = TRUE),2], ylab = "PC2 Loadings")
abline(h=0.035, col = "red")
dev.off()

jpeg(file="../../SupplementalFiles/FigureS3d.jpg", width=10, height=8, units = "in", res = 300)
plot(d2_less[order(d2_less$PC2loadings, decreasing = TRUE),2], ylab = "PC2 Loadings")
abline(h=-0.035, col = "red")
dev.off()


#make list of proteins with temperature PC1 loadings values >= 0.025
cutd1 <- data.frame(d1[which(abs(d1$PC1loadings) >= 0.035),1])
colnames(cutd1) <- "protein"
cutd2 <- data.frame(d2[which(abs(d2$PC2loadings) >= 0.035),1])
colnames(cutd2) <- "protein"
cutd12 <- unique(rbind(cutd1,cutd2))
```

number of proteins affected by time at loadings value > 0.025 or < -0.025
```{r, echo = FALSE}
nrow(cutd12)
```


**heatmap of proteins affected by time based on time effect PC1 and PC2 loadings value cutoffs**
```{r avgNSAF_timeEffectPC1and2_cutoff0.025_heatmap_OrderedByTime, echo = FALSE, fig.height=10, fig.width=10, echo = FALSE}
#make a list of cutoff proteins with normalized abundance
cut12_data <-data[,which(colnames(data) %in% cutd12$protein)]
rownames(cut12_data) <- paste(data$dpf, data$temp, sep="_")

#make data frames to store heatmap color values
#columns are dpf/temp; color 23C blue, 29C red, each dpf a different shade of gray
ColSideColors<- cbind(Temp=c("brown",rep(c("cyan3","magenta2"),6)))

hm <- as.dist(1-cor(cut12_data, use="pa"))
hm2 <- hclust(hm, method = 'complete')
plot(hm2, cex = 0.3)
abline(h=1.5, col = "red")

#Identify clades in heatmap
# define some clusters

#make a list of proteins with cluster IDs based on where the tree is cut
mycl <- cutree(hm2, h=1.5)

#assign colors to cluster IDs
clusterCols <- colorRamps::primary.colors(length(unique(mycl)))
myClusterSideBar <- clusterCols[mycl]


#plot heat map 
jpeg(file="../../Figures/Figure3b.jpg", width=8, height=8, units = "in", res = 300)

heatmap3(t.data.frame(cut12_data),cexCol = 1.5,RowSideColors=myClusterSideBar,ColSideColors = ColSideColors,RowAxisColors=1,ColAxisColors=1,hclustfun=hclust, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), method = "complete", scale = "row", Colv=NA,labRow = NA, labCol = c(19,21,21,23,23,25,25,27,27,29,29,31,31), RowSideLabs = NA, ColSideLabs = NA)
dev.off()

#get color name
clusterColor <- lapply(myClusterSideBar, color.id)
clusterColor <- lapply(clusterColor, `[[`,1)
clusterColor <- data.frame(unlist(clusterColor), stringsAsFactors = FALSE)

foo <- cbind(data.frame(mycl), clusterColor)
#add protein column to data frame for merging
foo$protein_ID <- rownames(foo)

#add protein column back to NSAF data frame for merging
cut12_data_proteinID <- data.frame(t(cut12_data))
cut12_data_proteinID$protein_ID <- rownames(cut12_data_proteinID)

clade.prot <- merge(foo,cut12_data_proteinID,  by = "protein_ID")

colnames(clade.prot)[3] <- "ClusterColor"
#simplify clade names
clade.prot$SimpleClusterColor <- clade.prot$ClusterColor
clade.prot$SimpleClusterColor <- gsub("darkslategray1","light blue", clade.prot$SimpleClusterColor)
clade.prot$SimpleClusterColor <- gsub("gray50","gray", clade.prot$SimpleClusterColor)


#remove day 0 from clade prot
clade.prot_noD0 <- clade.prot[,-4]

#normalize within terms (autoscale by row)
#calculate row means and SD
clade.prot_noD0$mean <- apply(clade.prot_noD0[,4:15], 1, mean)
clade.prot_noD0$sd <- apply(clade.prot_noD0[,4:15],1, sd)

#sweep second argument is margin which corresponds to 1 (rows) or 2 (columns); this is for autoscaling by row to make colors relative to one another like the heatmap does
#subtract the row mean from each value
clade.prot_no_D0_norm <- sweep(clade.prot_noD0[,4:15],1,clade.prot_noD0$mean)
#divide each value by the row standard deviation
clade.prot_no_D0_norm <- sweep(clade.prot_no_D0_norm[,1:12],1,clade.prot_noD0$sd, FUN = "/")

clade.prot_no_D0_norm <- cbind(clade.prot_noD0[,c("protein_ID", "SimpleClusterColor","ClusterColor")], clade.prot_no_D0_norm)

STACKED_NSAF <- tidyr::gather(clade.prot_no_D0_norm, dpftemp, NSAF, 4:15)
STACKED_NSAF$dpf <- gsub("X|_.*","",STACKED_NSAF$dpftemp)
STACKED_NSAF$temp <- as.factor(gsub(".*_","",STACKED_NSAF$dpftemp))
str(STACKED_NSAF)
STACKED_NSAF$dpf <- as.integer(STACKED_NSAF$dpf)

#make mean summary table for plotting summary lines of protein abundances/clade
STACKED_NSAF_mean <- STACKED_NSAF %>%
    dplyr::group_by(dpf, temp, SimpleClusterColor,ClusterColor) %>%
    dplyr::summarise(n = n(), mean=mean(NSAF), median=median(NSAF), sd=sd(NSAF)) %>%
    dplyr::mutate(sem = sd / sqrt(n - 1),
      CI_lower = mean + qt((1-0.95)/2, n - 1) * sem,
      CI_upper = mean - qt((1-0.95)/2, n - 1) * sem)

jpeg(file="../../Figures/Figure3c.jpg", width=25, height=5, units = "in", res = 300)
ggplot(STACKED_NSAF_mean, aes(x=dpf, y=mean, color = temp)) + geom_line(aes(x=dpf, y=mean, color=temp)) + xlab("dpf") + ylab(" autoscaled NSAF values") + geom_ribbon(aes(ymin=CI_lower,ymax=CI_upper,fill=temp),color="grey70",alpha=0.4)  + facet_wrap(~factor(SimpleClusterColor, levels = c("light blue","purple","green","gray","black")), ncol = 5) + scale_color_manual(values=c("cyan3", "magenta2")) + scale_fill_manual(values=c("cyan3", "magenta2")) + scale_x_continuous(breaks = c(21,23,25,27,29,31), labels = c(21,23,25,27,29,31)) + theme_bw() + theme(text = element_text(size=30),strip.text = element_text(face = "bold",size = 30),strip.background = element_rect(fill="white"))
dev.off()

#create a simple plot for common trend over time 
STACKED_NSAF_mean_dpf <- STACKED_NSAF %>%
    dplyr::group_by(dpf, SimpleClusterColor,ClusterColor) %>%
    dplyr::summarise(n = n(), mean=mean(NSAF), median=median(NSAF), sd=sd(NSAF)) %>%
    dplyr::mutate(sem = sd / sqrt(n - 1),
      CI_lower = mean + qt((1-0.95)/2, n - 1) * sem,
      CI_upper = mean - qt((1-0.95)/2, n - 1) * sem)

jpeg("../../Figures/Figure4_colLabs.jpg", width = 5, height = 1, units = "in", res =300)
ggplot(STACKED_NSAF_mean_dpf, aes(x=dpf, y=mean)) + geom_line(aes(x=dpf, y=mean), size = 1) +ylab("mean autoscaled NSAF") + facet_wrap(~factor(SimpleClusterColor, levels = c("light blue","purple","green","gray","black")), ncol = 5)  + scale_x_continuous(breaks = c(21,23,25,27,29,31), labels = c(21,23,25,27,29,31)) + theme_bw()+ theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(),axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),strip.text = element_text(face = "bold",size = 15),strip.background = element_rect(fill="white")) 
dev.off()
```

Format table with all proteins and their loadings values for PC1 and PC2
```{r}
#PC1 and PC2 loadings for ASCA$1$svd$[,1:2]
time_ASCA_PC1_2_loadings <- cbind(protnames, ASCA$`1`$svd$v[,1:2])
colnames(time_ASCA_PC1_2_loadings) <- c("protein_ID", "ASCA_PC1_loadings", "ASCA_PC2_loadings")

time_ASCA_PC1_2_loadings$ASCA_PC_threshold_pass <- "No"

for(i in 1:nrow(time_ASCA_PC1_2_loadings)){
  if(time_ASCA_PC1_2_loadings[i,"protein_ID"] %in% cutd12$protein){
    time_ASCA_PC1_2_loadings$ASCA_PC_threshold_pass[i] <- "Yes"
    }
  else(time_ASCA_PC1_2_loadings$ASCA_PC_threshold_pass[i] <- "No")
}

#Merge cluster color and id with PC loadings data
time_ASCA_PC1_2_loadings <- merge(time_ASCA_PC1_2_loadings, clade.prot[,c("protein_ID", "ClusterColor", "SimpleClusterColor")], by = "protein_ID", all.x = TRUE)
```


## FUNCTIONAL ANALYSIS


**read in and format uniprot data**
```{r}
#read in uniprot mapping
uniprot <- read.csv("../../Data/all_giga-uniprot-blastP-out.nopipe.annotations.tab", sep ="\t", header = FALSE, stringsAsFactors = FALSE)
#rename uniprot columns
colnames(uniprot) <- c("protein_ID","Entry", "Entry_name", "perc_ident_match", "align_len", "num_mismatch", "num_gaps","querStart", "querEnd", "subjStart", "subjEnd", "evalue", "bitscore","Entry.1","Entry_name.1", "Protein_names", "Gene_names", "Organism", "Protein_length","Pathway", "GO_bp", "GO","GO_IDs", "Protein_fams")
#convert columns that should be numeric to numeric
str(uniprot)
uniprot[,4:13] <- lapply(uniprot[,4:13], as.numeric)
#NAs introduced by coercion
#this is because the entries that are unmappaed have NA values
str(uniprot)

#filter for choyp proteins and for evalue <= 10x10^-10 (unmapped proteins get removed here)
uniprot_filtered <- uniprot[grep("CHOYP",uniprot$protein_ID),]
uniprot_filtered <- uniprot[which(uniprot$evalue <= 1*10^-10),]


```

**Match GO terms to sig protein lists**

```{r}
#Get GO IDs for silo 3 and 9 sig. proteins
sig_pro_GO <- merge(time_ASCA_PC1_2_loadings[which(time_ASCA_PC1_2_loadings$ASCA_PC_threshold_pass == "Yes"),c("protein_ID", "ClusterColor")], uniprot_filtered[,c("protein_ID","GO_IDs")], by ="protein_ID")

#count number of proteins with GO terms
nrow(sig_pro_GO)

split_terms <- strsplit(as.character(sig_pro_GO$GO_IDs), ";") #split into multiple GO ids
STACKED_sig_pro_GO <- data.frame(protein_ID = rep.int(sig_pro_GO$protein_ID, sapply(split_terms, length)), GO = unlist(split_terms)) #list all genes with each of their GO terms in a single row 
```

**MATCH GO TERMS TO BACKGROUND LIST**
```{r}
#Get GO IDs for silo 3 and 9 ALL. proteins
ALL_pro_GO <-uniprot_filtered[which(uniprot_filtered$protein_ID %in% colnames(data)),c("protein_ID","GO_IDs")]

#count number of proteins with GO terms
nrow(ALL_pro_GO)

split_terms <- strsplit(as.character(ALL_pro_GO$GO_IDs), ";") #split into multiple GO ids
STACKED_ALL_pro_GO <- data.frame(protein_ID = rep.int(ALL_pro_GO$protein_ID, sapply(split_terms, length)), GO = unlist(split_terms)) #list all genes with each of their GO terms in a single row 
```

### TopGO 
```{r}

###this chunk generates p-values for weight01 fisher test like Liew paper (https://advances.sciencemag.org/content/4/6/eaar8028/tab-figures-data#ref-43)

#background list ALL_pro_GO
#sig pro list sig_pro_GO

topgo_bkgd <- ALL_pro_GO
topgo_bkgd$GO_IDs <- gsub(";",",",topgo_bkgd$GO_IDs)
write.table(topgo_bkgd, "TopGO_bkgd.txt", sep = "\t",quote = FALSE, row.names = FALSE, col.names = FALSE)

geneID2GO_bkgd <- readMappings(file = "TopGO_bkgd.txt")


#to switch between time and temp run line 66 in time and temp functional analysis scripts
topgo_sig <- sig_pro_GO
topgo_sig$GO_IDs <- gsub(";",",",topgo_sig$GO_IDs)

topgo_sig_1 <- topgo_sig[which(topgo_sig$ClusterColor == "darkslategray1"),c("protein_ID", "GO_IDs")]

topgo_sig_2 <- topgo_sig[which(topgo_sig$ClusterColor == "purple"),c("protein_ID", "GO_IDs")]

topgo_sig_3 <- topgo_sig[which(topgo_sig$ClusterColor == "green"),c("protein_ID", "GO_IDs")]

topgo_sig_4 <- topgo_sig[which(topgo_sig$ClusterColor == "gray50"),c("protein_ID", "GO_IDs")]

topgo_sig_5 <- topgo_sig[which(topgo_sig$ClusterColor == "black"),c("protein_ID", "GO_IDs")]

geneUniverse <- names(geneID2GO_bkgd) 

genesOfInterest_1 <- topgo_sig_1
genesOfInterest_1 <- as.character(genesOfInterest_1$protein_ID)

geneList_1 <- factor(as.integer(geneUniverse %in% genesOfInterest_1))
names(geneList_1) <- geneUniverse

myGOdata_1 <- new("topGOdata", description="time_clade1", ontology="BP", allGenes=geneList_1  ,annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_1 <- runTest(myGOdata_1, algorithm="weight01", statistic="fisher") 

allRes_1 <- GenTable(myGOdata_1, weight = resultFisher_1,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 2

genesOfInterest_2 <- topgo_sig_2
genesOfInterest_2 <- as.character(genesOfInterest_2$protein_ID)

geneList_2 <- factor(as.integer(geneUniverse %in% genesOfInterest_2))
names(geneList_2) <- geneUniverse

myGOdata_2 <- new("topGOdata", description="time_clade2", ontology="BP", allGenes=geneList_2,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_2 <- runTest(myGOdata_2, algorithm="weight01", statistic="fisher") 

allRes_2 <- GenTable(myGOdata_2, weight = resultFisher_2,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 3

genesOfInterest_3 <- topgo_sig_3
genesOfInterest_3 <- as.character(genesOfInterest_3$protein_ID)

geneList_3 <- factor(as.integer(geneUniverse %in% genesOfInterest_3))
names(geneList_3) <- geneUniverse

myGOdata_3 <- new("topGOdata", description="time_clade3", ontology="BP", allGenes=geneList_3,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_3 <- runTest(myGOdata_3, algorithm="weight01", statistic="fisher") 

allRes_3 <- GenTable(myGOdata_3, weight = resultFisher_3,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 4

genesOfInterest_4 <- topgo_sig_4
genesOfInterest_4 <- as.character(genesOfInterest_4$protein_ID)

geneList_4 <- factor(as.integer(geneUniverse %in% genesOfInterest_4))
names(geneList_4) <- geneUniverse

myGOdata_4 <- new("topGOdata", description="time_clade4", ontology="BP", allGenes=geneList_4,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_4 <- runTest(myGOdata_4, algorithm="weight01", statistic="fisher") 

allRes_4 <- GenTable(myGOdata_4, weight = resultFisher_4,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 5

genesOfInterest_5 <- topgo_sig_5
genesOfInterest_5 <- as.character(genesOfInterest_5$protein_ID)

geneList_5 <- factor(as.integer(geneUniverse %in% genesOfInterest_5))
names(geneList_5) <- geneUniverse

myGOdata_5 <- new("topGOdata", description="time_clade5", ontology="BP", allGenes=geneList_5,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_5 <- runTest(myGOdata_5, algorithm="weight01", statistic="fisher") 


allRes_5 <- GenTable(myGOdata_5, weight = resultFisher_5,orderBy = "weight", ranksOf = "classic", topNodes = 50)


sub_allRes1 <- allRes_1[which(allRes_1$Annotated >=5 & allRes_1$weight <= 0.05),]
sub_allRes2 <- allRes_2[which(allRes_2$Annotated >=5 & allRes_2$weight <= 0.05),]
sub_allRes3 <- allRes_3[which(allRes_3$Annotated >=5 & allRes_3$weight <= 0.05),]
sub_allRes4 <- allRes_4[which(allRes_4$Annotated >=5 & allRes_4$weight <= 0.05),]
sub_allRes5 <- allRes_5[which(allRes_5$Annotated >=5 & allRes_5$weight <= 0.05),]


#bind all result subsets into one data frame
sub_allRes <- rbind(sub_allRes1,sub_allRes2,sub_allRes3,sub_allRes4,sub_allRes5)

sub_allRes <- sub_allRes[,1:2]
sub_allRes <- unique(sub_allRes)

sub_allRes <- merge(sub_allRes, sub_allRes1[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[3] <- "clade1"

sub_allRes <- merge(sub_allRes, sub_allRes2[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[4] <- "clade2"

sub_allRes <- merge(sub_allRes, sub_allRes3[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[5] <- "clade3"

sub_allRes <- merge(sub_allRes, sub_allRes4[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[6] <- "clade4"

sub_allRes <- merge(sub_allRes, sub_allRes5[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[7] <- "clade5"
```

#create time GO enrichment table
```{r}
allRes_1$clade <- "light blue"
allRes_2$clade <- "purple"
allRes_3$clade <- "green"
allRes_4$clade <- "gray"
allRes_5$clade <- "black"
time_allRes <- rbind(allRes_1, allRes_2, allRes_3, allRes_4, allRes_5)
```

#get GO slims for clade terms
```{r}
#make list of unique GO terms without a protein ID column
GO_clade1 <- sub_allRes[which(!is.na(sub_allRes$clade1)),"GO.ID"]
GO_clade2 <- sub_allRes[which(!is.na(sub_allRes$clade2)),"GO.ID"]
GO_clade3 <- sub_allRes[which(!is.na(sub_allRes$clade3)),"GO.ID"]
GO_clade4 <- sub_allRes[which(!is.na(sub_allRes$clade4)),"GO.ID"]
GO_clade5 <- sub_allRes[which(!is.na(sub_allRes$clade5)),"GO.ID"]



#Use GSEA to generate list of GO Slim BP terms for each clade
#DF will have "term", "GOid", "GOcategory"

#goslims with GSEA
collection_clade1 <- GOCollection(GO_clade1)
collection_clade2 <- GOCollection(GO_clade2)
collection_clade3 <- GOCollection(GO_clade3)
collection_clade4 <- GOCollection(GO_clade4)
collection_clade5 <- GOCollection(GO_clade5)

#I downloaded goslim_generic.obo from http://geneontology.org/docs/go-subset-guide/
#then i moved it to the R library for GSEABase in the extdata folder
fl <- system.file("extdata", "goslim_generic.obo", package="GSEABase")
slim <- getOBOCollection(fl)
slims_clade1 <- data.frame(goSlim(collection_clade1, slim, "BP"))
slims_clade2 <- data.frame(goSlim(collection_clade2, slim, "BP"))
slims_clade3 <- data.frame(goSlim(collection_clade3, slim, "BP"))
slims_clade4 <- data.frame(goSlim(collection_clade4, slim, "BP"))
slims_clade5 <- data.frame(goSlim(collection_clade5, slim, "BP"))

#bind all clade slim terms together
slims <- rbind(slims_clade1, slims_clade2, slims_clade3, slims_clade4, slims_clade5)

slims <- data.frame(unique(slims$Term))
colnames(slims)[1] <- "Term"

slims <- merge(slims, slims_clade1[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[2] <- "clade1"

slims <- merge(slims, slims_clade2[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[3] <- "clade2"

slims <- merge(slims, slims_clade3[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[4] <- "clade3"

slims <- merge(slims, slims_clade4[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[5] <- "clade4"

slims <- merge(slims, slims_clade5[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[6] <- "clade5"

#remove rows without any counts, and biological_process row

slims <- slims[which(apply(slims[,-1],1,sum)!= 0 & slims$Term != "biological_process"),]

slims_m <- slims[,-1]
rownames(slims_m) <- slims$Term
slims_m <- data.matrix(slims_m)
colnames(slims_m) <- gsub("clade","",colnames(slims_m))

#replace truncated terms with full terms
rownames(slims_m) <- gsub("nucleobase-containing compound cata...","nucleobase-containing compound catabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("generation of precursor metabolites...","generation of precursor metabolites and energy", rownames(slims_m))

rownames(slims_m) <- gsub("cytoskeleton-dependent intracellula...","cytoskeleton-dependent intracellular transport", rownames(slims_m))

rownames(slims_m) <- gsub("cellular protein modification proce...","cellular protein modification process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic p", rownames(slims_m))

rownames(slims_m) <- gsub("cellular amino acid metabolic proce...","cellular amino acid metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("anatomical structure formation invo...","anatomical structure formation involved in morphogenesis", rownames(slims_m))

breaks <- c(seq(-0.5,12,by = 1))

jpeg(file="../../Figures/Figure4.jpg", width=8, height=9, units = "in", res = 300)
heatmap.2(slims_m,margins = c(5,27),cexRow = 1.2, Colv=NA,revC = TRUE,breaks = breaks,labCol = NA, ColSideColors = c("darkslategray1", "purple", "green", "gray50", "black"),density.info = "none", trace = "none",col = c("white", viridis(length(breaks)-2)), scale = "none",rowsep = 1:nrow(slims_m),colsep=1:ncol(slims_m),sepwidth=c(0.01,0.01),sepcolor="white", hclustfun = function(x) hclust(x, method="average"), distfun = function(x) as.dist(1-cor(t(x), use = "pa")))
dev.off()
```

Create table with GO temrs, slim terms, protein IDs, and gene IDs
```{r}
### make this table only for significantly enriched GO terms

#first define clade list
clust_list <- colnames(sub_allRes[,grep("clade", colnames(sub_allRes))])
#then assign cluster color to clade number
names(clust_list) <- c("darkslategray1", "purple", "green", "gray50", "black")

for(i in 1:nrow(topgo_sig)){
  clade_sig_GO <- na.omit(sub_allRes[,c("GO.ID",clust_list[[topgo_sig$ClusterColor[i]]])])
  A = unlist(strsplit(topgo_sig$GO_IDs[i], ", "))
  B = clade_sig_GO[,"GO.ID"]
  if(length(intersect(A,B)) >= 1){
    topgo_sig$sig_GO[i] <- toString(intersect(A,B))
  }
  else{
    topgo_sig$sig_GO[i] <- NA
  }
}

prot_slims <- data.frame()
for (i in 1:length(topgo_sig$protein_ID)){
  if(!is.na(topgo_sig$sig_GO[i])){
    GO_IDs <- unlist(strsplit(topgo_sig$sig_GO[i], ", "))
    collection_clade <- GOCollection(GO_IDs)
    fl <- system.file("extdata", "goslim_generic.obo", package="GSEABase")
    slim <- getOBOCollection(fl)
    slims_clade <- data.frame(goSlim(collection_clade, slim, "BP"))
    slims_clade$protein_ID <- topgo_sig$protein_ID[i]
    prot_slims <- rbind(prot_slims, slims_clade)
  }
}

prot_slims_summry <- prot_slims[which(prot_slims$Count !=0 & prot_slims$Term != "biological_process"),] %>% dplyr::group_by(protein_ID) %>% dplyr::summarise(GO_slim = toString(Term)) %>% ungroup()

#replace truncated terms with full terms
prot_slims_summry$GO_slim <- gsub("nucleobase-containing compound cata...","nucleobase-containing compound catabolic process", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("generation of precursor metabolites...","generation of precursor metabolites and energy", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cytoskeleton-dependent intracellula...","cytoskeleton-dependent intracellular transport", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cellular protein modification proce...","cellular protein modification process", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic process", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic p", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cellular amino acid metabolic proce...","cellular amino acid metabolic process", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("anatomical structure formation invo...","anatomical structure formation involved in morphogenesis", prot_slims_summry$GO_slim)

#replace , with ;
prot_slims_summry$GO_slim <- gsub(",",";", prot_slims_summry$GO_slim)


prot_slims_summry <- merge(topgo_sig,prot_slims_summry, by = "protein_ID", all.x = TRUE)

#merge with rest of uniprot data
prot_slims_summry_uni <- merge(uniprot_filtered[,c("protein_ID", "Protein_names", "Protein_fams", "Gene_names")],prot_slims_summry, by = "protein_ID")

#reorder columns
prot_slims_summry_uni <- data.frame(prot_slims_summry_uni[,c(1,5,2:4,6:8)])

prot_slims_summry_uni <- prot_slims_summry_uni[order(prot_slims_summry_uni$ClusterColor, prot_slims_summry_uni$GO_slim),]

colnames(prot_slims_summry_uni)[7] <- "sig_bp_GO"
colnames(prot_slims_summry_uni)[8] <- "sig_bp_GOslim"

prot_slims_summry_uni <- sapply(prot_slims_summry_uni, gsub, pattern=",", replacement=";") 
```

Make supplementary table 4
```{r}
time_ASCA_PC1_2_loadings_GO <- merge(time_ASCA_PC1_2_loadings,prot_slims_summry_uni, by = c("protein_ID", "ClusterColor"), all = TRUE)

time_ASCA_PC1_2_loadings_GO <- time_ASCA_PC1_2_loadings_GO[,c(1,3:5,2,6:12)]

#add full go terms and uniprot ID columns
time_ASCA_PC1_2_loadings_GO <- merge(time_ASCA_PC1_2_loadings_GO,uniprot_filtered[,c("protein_ID","Entry","GO")], by = "protein_ID", all.x = T)

#rename columns
colnames(time_ASCA_PC1_2_loadings_GO)[13:14] <- c("UniprotID", "GO_terms")

#reorder columns
time_ASCA_PC1_2_loadings_GO <- time_ASCA_PC1_2_loadings_GO[,c(1:6,13,7:9,14,10:12)]

#rename GO_ID column
colnames(time_ASCA_PC1_2_loadings_GO)[12] <- "GO_bp_IDs"


#reorder table by sig bp terms
time_ASCA_PC1_2_loadings_GO <- time_ASCA_PC1_2_loadings_GO[order(time_ASCA_PC1_2_loadings_GO$ASCA_PC_threshold_pass,time_ASCA_PC1_2_loadings_GO$ClusterColor,time_ASCA_PC1_2_loadings_GO$sig_bp_GO,time_ASCA_PC1_2_loadings_GO$GO_bp_IDs,time_ASCA_PC1_2_loadings_GO$Protein_names, decreasing = TRUE),]

#make sure there are no commas
time_ASCA_PC1_2_loadings_GO <- sapply(time_ASCA_PC1_2_loadings_GO, gsub, pattern=",", replacement=";")

write.csv(time_ASCA_PC1_2_loadings_GO, "../../SupplementalFiles/Supplemental_Table_4.csv", quote = FALSE, row.names = FALSE)

```

#######################################
Analysis of Temperature-affected proteins
######################################

**Match GO terms to sig protein lists**

```{r}
#Get GO IDs for silo 3 and 9 sig. proteins
sig_pro_GO <- merge(temp_ASCA_PC1_loadings[which(temp_ASCA_PC1_loadings$ASCA_PC_threshold_pass == "Yes"),c("protein_ID", "ClusterColor")], uniprot_filtered[,c("protein_ID","GO_IDs")], by ="protein_ID")

#count number of proteins with GO terms
nrow(sig_pro_GO)

split_terms <- strsplit(as.character(sig_pro_GO$GO_IDs), ";") #split into multiple GO ids
STACKED_sig_pro_GO <- data.frame(protein_ID = rep.int(sig_pro_GO$protein_ID, sapply(split_terms, length)), GO = unlist(split_terms)) #list all genes with each of their GO terms in a single row 
```

### TopGO 
```{r}

###this chunk generates p-values for weight01 fisher test like Liew paper (https://advances.sciencemag.org/content/4/6/eaar8028/tab-figures-data#ref-43)
topgo_sig <- sig_pro_GO
topgo_sig$GO_IDs <- gsub(";",",",topgo_sig$GO_IDs)

topgo_sig_1 <- topgo_sig[which(topgo_sig$ClusterColor == "cyan4"),c("protein_ID", "GO_IDs")]

topgo_sig_2 <- topgo_sig[which(topgo_sig$ClusterColor == "darkorange3"),c("protein_ID", "GO_IDs")]

geneUniverse <- names(geneID2GO_bkgd) 

genesOfInterest_1 <- topgo_sig_1
genesOfInterest_1 <- as.character(genesOfInterest_1$protein_ID)

geneList_1 <- factor(as.integer(geneUniverse %in% genesOfInterest_1))
names(geneList_1) <- geneUniverse

myGOdata_1 <- new("topGOdata", description="temp_clade1", ontology="BP", allGenes=geneList_1  ,annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_1 <- runTest(myGOdata_1, algorithm="weight01", statistic="fisher") 

allRes_1 <- GenTable(myGOdata_1, weight = resultFisher_1,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 2

genesOfInterest_2 <- topgo_sig_2
genesOfInterest_2 <- as.character(genesOfInterest_2$protein_ID)

geneList_2 <- factor(as.integer(geneUniverse %in% genesOfInterest_2))
names(geneList_2) <- geneUniverse

myGOdata_2 <- new("topGOdata", description="temp_clade2", ontology="BP", allGenes=geneList_2,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_2 <- runTest(myGOdata_2, algorithm="weight01", statistic="fisher") 

allRes_2 <- GenTable(myGOdata_2, weight = resultFisher_2,orderBy = "weight", ranksOf = "classic", topNodes = 50)


sub_allRes1 <- allRes_1[which(allRes_1$Annotated >=5 & allRes_1$weight <= 0.05),]
sub_allRes2 <- allRes_2[which(allRes_2$Annotated >=5 & allRes_2$weight <= 0.05),]

#bind all result subsets into one data frame
sub_allRes <- rbind(sub_allRes1,sub_allRes2)

sub_allRes <- sub_allRes[,1:2]
sub_allRes <- unique(sub_allRes)

sub_allRes <- merge(sub_allRes, sub_allRes1[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[3] <- "clade1"

sub_allRes <- merge(sub_allRes, sub_allRes2[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[4] <- "clade2"

```

#create temp GO enrichment table
```{r}
allRes_1$clade <- "dark teal"
allRes_2$clade <- "orange"
temp_allRes <- rbind(allRes_1, allRes_2)
```


#get GO slims for clade terms
```{r}
#make list of unique GO terms without a protein ID column
GO_clade1 <- sub_allRes[which(!is.na(sub_allRes$clade1)),"GO.ID"]
GO_clade2 <- sub_allRes[which(!is.na(sub_allRes$clade2)),"GO.ID"]

#Use GSEA to generate list of GO Slim BP terms for each clade
#DF will have "term", "GOid", "GOcategory"

#goslims with GSEA
collection_clade1 <- GOCollection(GO_clade1)
collection_clade2 <- GOCollection(GO_clade2)

#I downloaded goslim_generic.obo from http://geneontology.org/docs/go-subset-guide/
#then i moved it to the R library for GSEABase in the extdata folder
#make list of unique GO terms without a protein ID column
#moved it to the R library for GSEABase in the extdata folder
fl <- system.file("extdata", "goslim_generic.obo", package="GSEABase")
slim <- getOBOCollection(fl)
slims_clade1 <- data.frame(goSlim(collection_clade1, slim, "BP"))
slims_clade2 <- data.frame(goSlim(collection_clade2, slim, "BP"))

#bind all clade slim terms together
slims <- rbind(slims_clade1, slims_clade2)

slims <- data.frame(unique(slims$Term))
colnames(slims)[1] <- "Term"

slims <- merge(slims, slims_clade1[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[2] <- "clade1"

slims <- merge(slims, slims_clade2[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[3] <- "clade2"

#remove rows without any counts, and biological_process row

slims <- slims[which(apply(slims[,-1],1,sum)!= 0 & slims$Term != "biological_process"),]

slims_m <- slims[,-1]
rownames(slims_m) <- slims$Term
slims_m <- data.matrix(slims_m)
colnames(slims_m) <- gsub("clade","",colnames(slims_m))

#replace truncated terms with full terms
rownames(slims_m) <- gsub("nucleobase-containing compound cata...","nucleobase-containing compound catabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("generation of precursor metabolites...","generation of precursor metabolites and energy", rownames(slims_m))

rownames(slims_m) <- gsub("cytoskeleton-dependent intracellula...","cytoskeleton-dependent intracellular transport", rownames(slims_m))

rownames(slims_m) <- gsub("cellular protein modification proce...","cellular protein modification process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic p", rownames(slims_m))

rownames(slims_m) <- gsub("cellular amino acid metabolic proce...","cellular amino acid metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("anatomical structure formation invo...","anatomical structure formation involved in morphogenesis", rownames(slims_m))

breaks <- c(seq(-0.5,6,by = 1))

jpeg(file="../../Figures/Figure6.jpg", width=7, height=8, units = "in", res = 300)
heatmap.2(slims_m,margins = c(5,27),cexRow = 1.2, Colv=NA,revC = TRUE,breaks = breaks,labCol = NA, ColSideColors = c("cyan4", "darkorange3"),density.info = "none", trace = "none",col = c("white", viridis(length(breaks)-2)), scale = "none",rowsep = 1:nrow(slims_m),colsep=1:ncol(slims_m),sepwidth=c(0.01,0.01),sepcolor="white")
dev.off()



```

Create table with GO temrs, slim terms, protein IDs, and gene IDs
```{r}
### make this table only for significantly enriched GO terms

#first define clade list
clust_list <- colnames(sub_allRes[,grep("clade", colnames(sub_allRes))])
#then assign cluster color to clade number
names(clust_list) <- c("cyan4","darkorange3")

for(i in 1:nrow(topgo_sig)){
  clade_sig_GO <- na.omit(sub_allRes[,c("GO.ID",clust_list[[topgo_sig$ClusterColor[i]]])])
  A = unlist(strsplit(topgo_sig$GO_IDs[i], ", "))
  B = clade_sig_GO[,"GO.ID"]
  if(length(intersect(A,B)) >= 1){
    topgo_sig$sig_GO[i] <- toString(intersect(A,B))
  }
  else{
    topgo_sig$sig_GO[i] <- NA
  }
}

prot_slims <- data.frame()
for (i in 1:length(topgo_sig$protein_ID)){
  if(!is.na(topgo_sig$sig_GO[i])){
    GO_IDs <- unlist(strsplit(topgo_sig$sig_GO[i], ", "))
    collection_clade <- GOCollection(GO_IDs)
    fl <- system.file("extdata", "goslim_generic.obo", package="GSEABase")
    slim <- getOBOCollection(fl)
    slims_clade <- data.frame(goSlim(collection_clade, slim, "BP"))
    slims_clade$protein_ID <- topgo_sig$protein_ID[i]
    prot_slims <- rbind(prot_slims, slims_clade)
  }
}

prot_slims_summry <- prot_slims[which(prot_slims$Count !=0 & prot_slims$Term != "biological_process"),] %>% dplyr::group_by(protein_ID) %>% dplyr::summarise(GO_slim = toString(Term)) %>% ungroup()

#replace truncated terms with full terms
prot_slims_summry$GO_slim <- gsub("nucleobase-containing compound cata...","nucleobase-containing compound catabolic process", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("generation of precursor metabolites...","generation of precursor metabolites and energy", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cytoskeleton-dependent intracellula...","cytoskeleton-dependent intracellular transport", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cellular protein modification proce...","cellular protein modification process", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic process", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic p", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cellular amino acid metabolic proce...","cellular amino acid metabolic process", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("anatomical structure formation invo...","anatomical structure formation involved in morphogenesis", prot_slims_summry$GO_slim)

#replace , with ;
prot_slims_summry$GO_slim <- gsub(",",";", prot_slims_summry$GO_slim)


prot_slims_summry <- merge(topgo_sig,prot_slims_summry, by = "protein_ID", all.x = TRUE)

#merge with rest of uniprot data
prot_slims_summry_uni <- merge(uniprot_filtered[,c("protein_ID", "Protein_names", "Protein_fams", "Gene_names")],prot_slims_summry, by = "protein_ID")

#reorder columns
prot_slims_summry_uni <- data.frame(prot_slims_summry_uni[,c(1,5,2:4,6:8)])

prot_slims_summry_uni <- prot_slims_summry_uni[order(prot_slims_summry_uni$ClusterColor, prot_slims_summry_uni$GO_slim),]

colnames(prot_slims_summry_uni)[7] <- "sig_bp_GO"
colnames(prot_slims_summry_uni)[8] <- "sig_bp_GOslim"

prot_slims_summry_uni <- sapply(prot_slims_summry_uni, gsub, pattern=",", replacement=";") 

```

Make supplemental table 5
```{r}
temp_ASCA_PC1_loadings_GO <- merge(temp_ASCA_PC1_loadings,prot_slims_summry_uni, by = c("protein_ID", "ClusterColor"), all = TRUE)

temp_ASCA_PC1_loadings_GO <- temp_ASCA_PC1_loadings_GO[,c(1,3:4,2,5:ncol(temp_ASCA_PC1_loadings_GO))]

#add full go terms and uniprot ID columns
temp_ASCA_PC1_loadings_GO <- merge(temp_ASCA_PC1_loadings_GO,uniprot_filtered[,c("protein_ID","Entry","GO")], by = "protein_ID", all.x = T)

#rename columns
colnames(temp_ASCA_PC1_loadings_GO)[12:13] <- c("UniprotID", "GO_terms")

#reorder columns
temp_ASCA_PC1_loadings_GO <- temp_ASCA_PC1_loadings_GO[,c(1:5,12,6:8,13,9:11)]

#rename GO_ID column
colnames(temp_ASCA_PC1_loadings_GO)[11] <- "GO_bp_IDs"


#reorder table by sig bp terms
temp_ASCA_PC1_loadings_GO <- temp_ASCA_PC1_loadings_GO[order(temp_ASCA_PC1_loadings_GO$ASCA_PC_threshold_pass,temp_ASCA_PC1_loadings_GO$ClusterColor,temp_ASCA_PC1_loadings_GO$sig_bp_GO,temp_ASCA_PC1_loadings_GO$GO_bp_IDs,temp_ASCA_PC1_loadings_GO$Protein_names, decreasing = TRUE),]

#make sure there are no commas
temp_ASCA_PC1_loadings_GO <- sapply(temp_ASCA_PC1_loadings_GO, gsub, pattern=",", replacement=";")


write.csv(temp_ASCA_PC1_loadings_GO, "../../SupplementalFiles/Supplemental_Table_5.csv", quote = FALSE, row.names = FALSE)

```

#######################################
Analysis of Day21 and day27 affected proteins
######################################

**Match GO terms to proteins**

```{r}
#Get GO IDs for proteins
sig_pro_GO <- merge(PC_loadings[which(PC_loadings$PC_threshold_pass == "Yes" & PC_loadings$ClusterColor != "paleturquoise3"),c("protein_ID", "ClusterColor")], uniprot_filtered[,c("protein_ID","GO_IDs")], by ="protein_ID")

#count number of proteins with GO terms
nrow(sig_pro_GO)

split_terms <- strsplit(as.character(sig_pro_GO$GO_IDs), ";") #split into multiple GO ids
STACKED_sig_pro_GO <- data.frame(protein_ID = rep.int(sig_pro_GO$protein_ID, sapply(split_terms, length)), GO = unlist(split_terms)) #list all genes with each of their GO terms in a single row 

```

### TopGO 
```{r}

###this chunk generates p-values for weight01 fisher test like Liew paper (https://advances.sciencemag.org/content/4/6/eaar8028/tab-figures-data#ref-43)

#to switch between time and temp run line 66 in time and temp functional analysis scripts
topgo_sig <- sig_pro_GO
topgo_sig$GO_IDs <- gsub(";",",",topgo_sig$GO_IDs)

topgo_sig_1 <- topgo_sig[which(topgo_sig$ClusterColor == "wheat1"),c("protein_ID", "GO_IDs")]

topgo_sig_2 <- topgo_sig[which(topgo_sig$ClusterColor == "salmon"),c("protein_ID", "GO_IDs")]

topgo_sig_3 <- topgo_sig[which(topgo_sig$ClusterColor == "lightsteelblue"),c("protein_ID", "GO_IDs")]

geneUniverse <- names(geneID2GO_bkgd) 

genesOfInterest_1 <- topgo_sig_1
genesOfInterest_1 <- as.character(genesOfInterest_1$protein_ID)

geneList_1 <- factor(as.integer(geneUniverse %in% genesOfInterest_1))
names(geneList_1) <- geneUniverse

myGOdata_1 <- new("topGOdata", description="glob_clade1", ontology="BP", allGenes=geneList_1  ,annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_1 <- runTest(myGOdata_1, algorithm="weight01", statistic="fisher") 

allRes_1 <- GenTable(myGOdata_1, weight = resultFisher_1,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 2

genesOfInterest_2 <- topgo_sig_2
genesOfInterest_2 <- as.character(genesOfInterest_2$protein_ID)

geneList_2 <- factor(as.integer(geneUniverse %in% genesOfInterest_2))
names(geneList_2) <- geneUniverse

myGOdata_2 <- new("topGOdata", description="glob_clade2", ontology="BP", allGenes=geneList_2,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_2 <- runTest(myGOdata_2, algorithm="weight01", statistic="fisher") 

allRes_2 <- GenTable(myGOdata_2, weight = resultFisher_2,orderBy = "weight", ranksOf = "classic", topNodes = 50)

#clade 3

genesOfInterest_3 <- topgo_sig_3
genesOfInterest_3 <- as.character(genesOfInterest_3$protein_ID)

geneList_3 <- factor(as.integer(geneUniverse %in% genesOfInterest_3))
names(geneList_3) <- geneUniverse

myGOdata_3 <- new("topGOdata", description="glob_clade3", ontology="BP", allGenes=geneList_3,  annot = annFUN.gene2GO, gene2GO = geneID2GO_bkgd)

resultFisher_3 <- runTest(myGOdata_3, algorithm="weight01", statistic="fisher") 

allRes_3 <- GenTable(myGOdata_3, weight = resultFisher_3,orderBy = "weight", ranksOf = "classic", topNodes = 50)


sub_allRes1 <- allRes_1[which(allRes_1$Annotated >=5 & allRes_1$weight <= 0.05),]
sub_allRes2 <- allRes_2[which(allRes_2$Annotated >=5 & allRes_2$weight <= 0.05),]
sub_allRes3 <- allRes_3[which(allRes_3$Annotated >=5 & allRes_3$weight <= 0.05),]

#bind all result subsets into one data frame
sub_allRes <- rbind(sub_allRes1,sub_allRes2, sub_allRes3)

sub_allRes <- sub_allRes[,1:2]
sub_allRes <- unique(sub_allRes)

sub_allRes <- merge(sub_allRes, sub_allRes1[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[3] <- "clade1"

sub_allRes <- merge(sub_allRes, sub_allRes2[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[4] <- "clade2"

sub_allRes <- merge(sub_allRes, sub_allRes3[,c(1,6)], by = "GO.ID", all = TRUE)
colnames(sub_allRes)[5] <- "clade3"
```


#create temp GO enrichment table
```{r}
allRes_1$clade <- "pale yellow"
allRes_2$clade <- "salmon"
allRes_3$clade <- "pale purple"
d21_27_allRes <- rbind(allRes_1, allRes_2, allRes_3)
```

#output GO enrichment table for all clades
```{r}
#bind all clade GO enrichment
allRes <- rbind(time_allRes,temp_allRes, d21_27_allRes)
#reorder columns so that clade is first
allRes <- allRes[,c(ncol(allRes), 1:(ncol(allRes)-1))]
#substitute commas for semicolons
allRes <- sapply(allRes, gsub, pattern=",", replacement=";") 

#write out table
write.csv(allRes,"topGOenrichment_allClades.csv", row.names = F, quote = F)

```


#get GO slims for clade terms
```{r}

GO_clade1 <- sub_allRes[which(!is.na(sub_allRes$clade1)),"GO.ID"]
GO_clade2 <- sub_allRes[which(!is.na(sub_allRes$clade2)),"GO.ID"]
GO_clade3 <- sub_allRes[which(!is.na(sub_allRes$clade3)),"GO.ID"]

#Use GSEA to generate list of GO Slim BP terms for each clade
#DF will have "term", "GOid", "GOcategory"

#goslims with GSEA
collection_clade1 <- GOCollection(GO_clade1)
collection_clade2 <- GOCollection(GO_clade2)
collection_clade3 <- GOCollection(GO_clade3)

#make list of unique GO terms without a protein ID column
#moved it to the R library for GSEABase in the extdata folder
fl <- system.file("extdata", "goslim_generic.obo", package="GSEABase")
slim <- getOBOCollection(fl)
slims_clade1 <- data.frame(goSlim(collection_clade1, slim, "BP"))
slims_clade2 <- data.frame(goSlim(collection_clade2, slim, "BP"))
slims_clade3 <- data.frame(goSlim(collection_clade3, slim, "BP"))

#bind all clade slim terms together
slims <- rbind(slims_clade1, slims_clade2, slims_clade3)

slims <- data.frame(unique(slims$Term))
colnames(slims)[1] <- "Term"

slims <- merge(slims, slims_clade1[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[2] <- "clade1"

slims <- merge(slims, slims_clade2[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[3] <- "clade2"

slims <- merge(slims, slims_clade3[,c("Term", "Count")], by = "Term", all = TRUE)
colnames(slims)[4] <- "clade3"

#remove rows without any counts, and biological_process row

slims <- slims[which(apply(slims[,-1],1,sum)!= 0 & slims$Term != "biological_process"),]


slims_m <- slims[,-1]
rownames(slims_m) <- slims$Term
slims_m <- data.matrix(slims_m)
colnames(slims_m) <- gsub("clade","",colnames(slims_m))

#replace truncated terms with full terms
rownames(slims_m) <- gsub("nucleobase-containing compound cata...","nucleobase-containing compound catabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("generation of precursor metabolites...","generation of precursor metabolites and energy", rownames(slims_m))

rownames(slims_m) <- gsub("cytoskeleton-dependent intracellula...","cytoskeleton-dependent intracellular transport", rownames(slims_m))

rownames(slims_m) <- gsub("cellular protein modification proce...","cellular protein modification process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic p", rownames(slims_m))

rownames(slims_m) <- gsub("cellular amino acid metabolic proce...","cellular amino acid metabolic process", rownames(slims_m))

rownames(slims_m) <- gsub("anatomical structure formation invo...","anatomical structure formation involved in morphogenesis", rownames(slims_m))

breaks <- c(seq(-0.5,8.5, by = 1))

jpeg(file="../../Figures/Figure2C.jpg", width=7, height=7, units = "in", res = 300)
heatmap.2(slims_m,margins = c(5,27),cexRow = 1.2, Colv=NA,revC = TRUE,labCol = NA, breaks = breaks, ColSideColors = c("wheat1", "salmon", "lightsteelblue"),density.info = "none", trace = "none",col = c("white", viridis(length(breaks)-2)), scale = "none",rowsep = 1:nrow(slims_m),colsep=1:ncol(slims_m),sepwidth=c(0.01,0.01),sepcolor="white")
dev.off()
```

Create table with GO temrs, slim terms, protein IDs, and gene IDs
```{r}
#first define clade list
clust_list <- colnames(sub_allRes[,grep("clade", colnames(sub_allRes))])
#then assign cluster color to clade number
names(clust_list) <- c("wheat1","salmon", "lightsteelblue")

for(i in 1:nrow(topgo_sig)){
  clade_sig_GO <- na.omit(sub_allRes[,c("GO.ID",clust_list[[topgo_sig$ClusterColor[i]]])])
  A = unlist(strsplit(topgo_sig$GO_IDs[i], ", "))
  B = clade_sig_GO[,"GO.ID"]
  if(length(intersect(A,B)) >= 1){
    topgo_sig$sig_GO[i] <- toString(intersect(A,B))
  }
  else{
    topgo_sig$sig_GO[i] <- NA
  }
}


prot_slims <- data.frame()
for (i in 1:length(topgo_sig$protein_ID)){
  if(!is.na(topgo_sig$sig_GO[i])){
    GO_IDs <- unlist(strsplit(topgo_sig$sig_GO[i], ", "))
    collection_clade <- GOCollection(GO_IDs)
    fl <- system.file("extdata", "goslim_generic.obo", package="GSEABase")
    slim <- getOBOCollection(fl)
    slims_clade <- data.frame(goSlim(collection_clade, slim, "BP"))
    slims_clade$protein_ID <- topgo_sig$protein_ID[i]
    prot_slims <- rbind(prot_slims, slims_clade)
  }
}

prot_slims_summry <- prot_slims[which(prot_slims$Count !=0 & prot_slims$Term != "biological_process"),] %>% dplyr::group_by(protein_ID) %>% dplyr::summarise(GO_slim = toString(Term)) %>% ungroup()

#replace truncated terms with full terms
prot_slims_summry$GO_slim <- gsub("nucleobase-containing compound cata...","nucleobase-containing compound catabolic process", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("generation of precursor metabolites...","generation of precursor metabolites and energy", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cytoskeleton-dependent intracellula...","cytoskeleton-dependent intracellular transport", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cellular protein modification proce...","cellular protein modification process", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic process", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cellular nitrogen compound metaboli...","cellular nitrogen compound metabolic p", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("cellular amino acid metabolic proce...","cellular amino acid metabolic process", prot_slims_summry$GO_slim)

prot_slims_summry$GO_slim <- gsub("anatomical structure formation invo...","anatomical structure formation involved in morphogenesis", prot_slims_summry$GO_slim)

#replace , with ;
prot_slims_summry$GO_slim <- gsub(",",";", prot_slims_summry$GO_slim)


prot_slims_summry <- merge(topgo_sig,prot_slims_summry, by = "protein_ID", all.x = TRUE)

#merge with rest of uniprot data
prot_slims_summry_uni <- merge(uniprot_filtered[,c("protein_ID", "Protein_names", "Protein_fams", "Gene_names")],prot_slims_summry, by = "protein_ID")

#reorder columns
prot_slims_summry_uni <- data.frame(prot_slims_summry_uni[,c(1,5,2:4,6:8)])

prot_slims_summry_uni <- prot_slims_summry_uni[order(prot_slims_summry_uni$ClusterColor, prot_slims_summry_uni$GO_slim),]

colnames(prot_slims_summry_uni)[7] <- "sig_bp_GO"
colnames(prot_slims_summry_uni)[8] <- "sig_bp_GOslim"

prot_slims_summry_uni <- sapply(prot_slims_summry_uni, gsub, pattern=",", replacement=";") 
```

Make supplementary table 3
```{r}
PC_loadings_GO <- merge(PC_loadings,prot_slims_summry_uni, by = c("protein_ID", "ClusterColor"), all = TRUE)

PC_loadings_GO <- PC_loadings_GO[,c(1,3:5,2,6:ncol(PC_loadings_GO))]

#add full go terms and uniprot ID columns
PC_loadings_GO <- merge(PC_loadings_GO,uniprot_filtered[,c("protein_ID","Entry","GO")], by = "protein_ID", all.x = T)

#rename columns
colnames(PC_loadings_GO)[13:14] <- c("UniprotID", "GO_terms")

#reorder columns
PC_loadings_GO <- PC_loadings_GO[,c(1:6,13,7:9,14,10:12)]

#rename GO_ID column
colnames(PC_loadings_GO)[12] <- "GO_bp_IDs"

PC_loadings_GO <- PC_loadings_GO[order(PC_loadings_GO$PC_threshold_pass,PC_loadings_GO$ClusterColor,PC_loadings_GO$sig_bp_GO,PC_loadings_GO$GO_bp_IDs,PC_loadings_GO$Protein_names, decreasing = TRUE),]

#make sure there are no commas
PC_loadings_GO <- sapply(PC_loadings_GO, gsub, pattern=",", replacement=";")


write.csv(PC_loadings_GO, "../../SupplementalFiles/Supplemental_Table_3.csv", quote = FALSE, row.names = FALSE)

```


