---
title: "General_proteome_characterization"
author: "Shelly Trigg"
date: "4/3/2019"
output: rmarkdown::github_document
---

Load libraries
```{r}
library(tidyr)
library(GSEABase)
library(plyr)
library(RColorBrewer)
library(ggplot2)
```

read in data
```{r}
NSAF <- read.csv("~/Documents/GitHub/OysterSeedProject/analysis/nmds_R/silo3and9_nozerovals_AVGs.csv", stringsAsFactors = FALSE)
sr_lab_goslim <- read.csv("~/Documents/GitHub/OysterSeedProject/raw_data/background/GOSlim_terms.csv", stringsAsFactors = FALSE)
colnames(sr_lab_goslim)[1] <- "GO"
#For reading in uniprot, gannet must be mounted on the computer
uniprot <- read.csv("/Volumes/web/metacarcinus/Cgigas/all_giga-uniprot-blastP-out.nopipe.annotations.tab", sep ="\t", header = FALSE, stringsAsFactors = FALSE)
#rename uniprot columns
colnames(uniprot) <- c("protein_ID","Entry", "Entry_name", "perc_ident_match", "align_len", "num_mismatch", "num_gaps","querStart", "querEnd", "subjStart", "subjEnd", "evalue", "bitscore","Entry.1","Entry_name.1", "Protein_names", "Gene_names", "Organism", "Protein_length","Pathway", "GO_bp", "GO","GO_IDs", "Protein_fams")
```

I had already changed the zeros to 0.1, so I will sub that back with zero for this analysis
```{r}
#I had already changed the zeros to 0.1, so I will sub that back with zero for this analysis
NSAF[NSAF == 0.1] <- 0
```

separate 3 and 9
```{r}
#separate 3 and 9

silo3 <- NSAF[grep("23", NSAF$temp),]
silo9 <- NSAF[grep("29", NSAF$temp),]
```

reformat to long list
```{r}
silo3_t <- data.frame(t(silo3[,-c(1,2)]))
colnames(silo3_t) <- paste0("D",silo3$day)

silo9_t <- data.frame(t(silo9[,-c(1,2)]))
colnames(silo9_t) <- paste0("D",silo9$day)

```

remove proteins that were only in silo9
```{r}
no_val_proteins <- rownames(silo3_t[which(apply(silo3_t, 1, var) == 0),])
silo3_t_nozero <- silo3_t[-c(which(rownames(silo3_t) %in% no_val_proteins)),]
#confirm it worked
nrow(silo3_t)
nrow(silo3_t_nozero)
nrow(silo3_t)-nrow(silo3_t_nozero)
length(no_val_proteins)

no_val_proteins <- rownames(silo9_t[which(apply(silo9_t, 1, var) == 0),])
silo9_t_nozero <- silo9_t[-c(which(rownames(silo9_t) %in% no_val_proteins)),]
#confirm it worked
nrow(silo9_t)
nrow(silo9_t_nozero)
nrow(silo9_t)-nrow(silo9_t_nozero)
length(no_val_proteins)

```

add protein column
```{r}
silo3_t_nozero$protein_ID <- rownames(silo3_t_nozero)
silo3_t_nozero <- silo3_t_nozero[grep("CHOYP", silo3_t_nozero$protein_ID),]

silo9_t_nozero$protein_ID <- rownames(silo9_t_nozero)
silo9_t_nozero <- silo9_t_nozero[grep("CHOYP", silo9_t_nozero$protein_ID),]

```

merge with uniprot
```{r}
#combine uniprot and NSAF data
silo3_t_nozero_unip<- merge(silo3_t_nozero, uniprot, by = "protein_ID", all.x = TRUE)

silo9_t_nozero_unip<- merge(silo9_t_nozero, uniprot, by = "protein_ID", all.x = TRUE)
```

exclude proteins that didn't map to uniprot DB
```{r}
silo3_t_nozero_unip_mapped <- silo3_t_nozero_unip[-grep("unmapped", silo3_t_nozero_unip$Entry),]
nrow(silo3_t_nozero_unip_mapped)
silo3_t_nozero_unip_mapped <- silo3_t_nozero_unip_mapped[which(silo3_t_nozero_unip_mapped$evalue <= 10^-10),]
nrow(silo3_t_nozero_unip_mapped)

silo9_t_nozero_unip_mapped <- silo9_t_nozero_unip[-grep("unmapped", silo9_t_nozero_unip$Entry),]
nrow(silo9_t_nozero_unip_mapped)
silo9_t_nozero_unip_mapped <- silo9_t_nozero_unip_mapped[which(silo9_t_nozero_unip_mapped$evalue <= 10^-10),]
nrow(silo9_t_nozero_unip_mapped)

```

get go terms for each day for silo 3 (23C)
```{r}
#create empty data frames to fill in with loop
silo3_all_srlab_terms_per_day <- data.frame()
silo3_all_srlab_BPterms_per_day <- data.frame()

#create list of days
days <- colnames(silo3_t_nozero_unip_mapped[,2:7])

#loop through days and pull out GO IDs associated with proteins from specific day
for(i in days){
  silo3_pro_GO <- silo3_t_nozero_unip_mapped[which(silo3_t_nozero_unip_mapped[,i] != 0),c("protein_ID","GO_IDs")] # create day specific list of proteins and their associated GO IDs 

  silo3_pro_GOid_term <- data.frame() #create another empty data frame
#loop through each line of data frame containing one column with "protein_ID" and one column with list of "GO_IDs" separated by ";"
  for (j in 1:nrow(silo3_pro_GO)){
    #create a row that lists all GO IDs associated with one protein listed across rows, where each different GO ID gets put in it's own column
    silo3_pro_GOid_term_row <- data.frame(t(data.frame(strsplit(as.character(silo3_pro_GO$GO_IDs[j]),'; ', fixed = TRUE))), stringsAsFactors = FALSE)
    #add each row created in the line above to the empty data frame
    #this will add NAs in columns where less terms exist; e.g. if a protein only has two terms and another has 10, there will be 8 NAs added to the row for the protein with 2 terms
    silo3_pro_GOid_term <- rbind.fill(silo3_pro_GOid_term,silo3_pro_GOid_term_row)
  }

  #add protein IDs back to GO IDs
  silo3_pro_GOid_term <- cbind(silo3_pro_GO[,"protein_ID"], silo3_pro_GOid_term)
  #this results in a table with protein ID listed in one column and each next column contains a GO ID that was listed with the protein in Uniprot DB.
  
  #reshape data so that all GO ID columns are gathered in one column called "GO" 
  STACKED_silo3_pro_GOid_term <- tidyr::gather(silo3_pro_GOid_term,"protein_ID","GO", 2:ncol(silo3_pro_GOid_term))
  #exlude middle column which just contains the string "protein_ID" in each row
  STACKED_silo3_pro_GOid_term <- STACKED_silo3_pro_GOid_term[,c(1,3)]
  #remove duplicate rows
  STACKED_silo3_pro_GOid_term <- unique(STACKED_silo3_pro_GOid_term)
  colnames(STACKED_silo3_pro_GOid_term)[1] <- "protein_ID"
  #remove any rows where GO column has NA value. 
  STACKED_silo3_pro_GOid_term <- STACKED_silo3_pro_GOid_term[which(!is.na(STACKED_silo3_pro_GOid_term$GO)),]
  #this resulting data frame has two columns "protein_ID" and "GO"
  
  ##get go slim terms for each day
  STACKED_silo3_pro_GOid_term_sr_labGOslim <- merge(STACKED_silo3_pro_GOid_term, sr_lab_goslim, by = "GO", all.x = TRUE)
  
  STACKED_silo3_pro_GOid_term_sr_labGOslim$GOSlim_bin <- ifelse(is.na(STACKED_silo3_pro_GOid_term_sr_labGOslim$GOSlim_bin),"unmapped",STACKED_silo3_pro_GOid_term_sr_labGOslim$GOSlim_bin)
  
  silo3_all_srlab_terms <- data.frame(table(STACKED_silo3_pro_GOid_term_sr_labGOslim$GOSlim_bin))
  silo3_all_srlab_terms$day <- i
  #View(silo3_all_srlab_terms)
  
  silo3_all_srlab_BPterms <- data.frame(table(STACKED_silo3_pro_GOid_term_sr_labGOslim[which(STACKED_silo3_pro_GOid_term_sr_labGOslim$aspect=="P"),"GOSlim_bin"]))
  silo3_all_srlab_BPterms$day <- i
    
  silo3_all_srlab_terms_per_day <- rbind(silo3_all_srlab_terms_per_day,silo3_all_srlab_terms)
  silo3_all_srlab_BPterms_per_day <- rbind(silo3_all_srlab_BPterms_per_day,silo3_all_srlab_BPterms)
}

```

get go terms for each day for silo 9 (29C)
```{r}
silo9_all_srlab_terms_per_day <- data.frame()
silo9_all_srlab_BPterms_per_day <- data.frame()

days <- colnames(silo9_t_nozero_unip_mapped[,2:7])
for(i in days){
  silo9_pro_GO <- silo9_t_nozero_unip_mapped[which(silo9_t_nozero_unip_mapped[,i] != 0),c("protein_ID","GO_IDs")]
  
  silo9_pro_GOid_term <- data.frame()
  #loop through each line of data frame containing one column with "protein_ID" and one column with list of "GO_IDs" separated by ";"
  for (j in 1:nrow(silo9_pro_GO)){
    #create a row that lists all GO IDs associated with one protein listed across rows, where each different GO ID gets put in it's own column
    silo9_pro_GOid_term_row <- data.frame(t(data.frame(strsplit(as.character(silo9_pro_GO$GO_IDs[j]),'; ', fixed = TRUE))), stringsAsFactors = FALSE)
    #add each row created in the line above to the empty data frame
    #this will add NAs in columns where less terms exist; e.g. if a protein only has two terms and another has 10, there will be 8 NAs added to the row for the protein with 2 terms
    silo9_pro_GOid_term <- rbind.fill(silo9_pro_GOid_term,silo9_pro_GOid_term_row)
  }
  
  #add protein IDs back to GO IDs
  silo9_pro_GOid_term <- cbind(silo9_pro_GO[,"protein_ID"], silo9_pro_GOid_term)
  #this results in a table with protein ID listed in one column and each next column contains a GO ID that was listed with the protein in Uniprot DB.
  
  #reshape data so that all GO ID columns are gathered in one column called "GO" 
  STACKED_silo9_pro_GOid_term <- tidyr::gather(silo9_pro_GOid_term,"protein_ID","GO", 2:ncol(silo9_pro_GOid_term))
  #exlude middle column which just contains the string "protein_ID" in each row
  STACKED_silo9_pro_GOid_term <- STACKED_silo9_pro_GOid_term[,c(1,3)]
  #remove duplicate rows
  STACKED_silo9_pro_GOid_term <- unique(STACKED_silo9_pro_GOid_term)
  colnames(STACKED_silo9_pro_GOid_term)[1] <- "protein_ID"
  #remove any rows where GO column has NA value. 
  STACKED_silo9_pro_GOid_term <- STACKED_silo9_pro_GOid_term[which(!is.na(STACKED_silo9_pro_GOid_term$GO)),]
  #this resulting data frame has two columns "protein_ID" and "GO"
  
  ##get go slim terms for each day
  STACKED_silo9_pro_GOid_term_sr_labGOslim <- merge(STACKED_silo9_pro_GOid_term, sr_lab_goslim, by = "GO", all.x = TRUE)
  
  STACKED_silo9_pro_GOid_term_sr_labGOslim$GOSlim_bin <- ifelse(is.na(STACKED_silo9_pro_GOid_term_sr_labGOslim$GOSlim_bin),"unmapped",STACKED_silo9_pro_GOid_term_sr_labGOslim$GOSlim_bin)
  
  silo9_all_srlab_terms <- data.frame(table(STACKED_silo9_pro_GOid_term_sr_labGOslim$GOSlim_bin))
  silo9_all_srlab_terms$day <- i
  #View(silo9_all_srlab_terms)
  
  silo9_all_srlab_BPterms <- data.frame(table(STACKED_silo9_pro_GOid_term_sr_labGOslim[which(STACKED_silo9_pro_GOid_term_sr_labGOslim$aspect=="P"),"GOSlim_bin"]))
  silo9_all_srlab_BPterms$day <- i
  
  silo9_all_srlab_terms_per_day <- rbind(silo9_all_srlab_terms_per_day,silo9_all_srlab_terms)
  silo9_all_srlab_BPterms_per_day <- rbind(silo9_all_srlab_BPterms_per_day,silo9_all_srlab_BPterms)
}
```

make pie chart for each day of sr lab GO slim terms for all mapped proteins from silo 3(23C)
```{r, 23C_pie_charts}
#for color palette expansion: https://www.r-bloggers.com/how-to-expand-color-palette-with-ggplot-and-rcolorbrewer/
#for pie chart: http://www.sthda.com/english/wiki/ggplot2-pie-chart-quick-start-guide-r-software-and-data-visualization


getPalette = colorRampPalette(brewer.pal(11, "Spectral"))
blank_theme <- theme_minimal()+
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_blank(),
    panel.grid=element_blank(),
    axis.ticks = element_blank(),
    plot.title=element_text(size=14, face="bold")
  )

#loop through days and create plot for each day
for(i in days){
  p <- ggplot(silo3_all_srlab_BPterms_per_day[which(silo3_all_srlab_BPterms_per_day$day== i),], aes(x = "", y = Freq, fill = Var1)) + geom_bar(stat = "identity") + coord_polar("y", start = 0) + scale_fill_manual(values = getPalette(nrow(silo3_all_srlab_BPterms_per_day[which(silo3_all_srlab_BPterms_per_day$day== i),]))) + blank_theme + theme(axis.text.x = element_blank()) + ggtitle(paste0("functional characterization of 23C ",i," proteome"))
  print(p)
}

```

```{r, 29C_pie_charts}
for(i in days){
  p <- ggplot(silo9_all_srlab_BPterms_per_day[which(silo9_all_srlab_BPterms_per_day$day== i),], aes(x = "", y = Freq, fill = Var1)) + geom_bar(stat = "identity") + coord_polar("y", start = 0) + scale_fill_manual(values = getPalette(nrow(silo9_all_srlab_BPterms_per_day[which(silo9_all_srlab_BPterms_per_day$day== i),]))) + blank_theme + theme(axis.text.x = element_blank()) + ggtitle(paste0("functional characterization of 29C ",i," proteome"))
  print(p)
}

```
    
    
compare GO slim frequencies between temperatures
```{r}
compare_terms <- merge(silo3_all_srlab_BPterms_per_day, silo9_all_srlab_BPterms_per_day, by = c("Var1", "day"))
colnames(compare_terms) <- c("GOslim", "day", "23C", "29C")
print(compare_terms)
```

## analysis of comprehensive proteomes (all days) for each temperature

get GO terms for comprehensive proteome from 23C exposured seed
```{r}
#Subset "protein_ID" column and "GO IDs" column containing list of GO_IDs separated by ";" from all_sig0.1_pro_logFC_pval 
silo3_pro_GO <- silo3_t_nozero_unip_mapped[,c("protein_ID","GO_IDs")]

silo3_pro_GOid_term <- data.frame()
#loop through each line of data frame containing one column with "protein_ID" and one column with list of "GO_IDs" separated by ";"
for (i in 1:nrow(silo3_pro_GO)){
  #create a row that lists all GO IDs associated with one protein listed across rows, where each different GO ID gets put in it's own column
  silo3_pro_GOid_term_row <- data.frame(t(data.frame(strsplit(as.character(silo3_pro_GO$GO_IDs[i]),'; ', fixed = TRUE))), stringsAsFactors = FALSE)
  #add each row created in the line above to the empty data frame
  #this will add NAs in columns where less terms exist; e.g. if a protein only has two terms and another has 10, there will be 8 NAs added to the row for the protein with 2 terms
  silo3_pro_GOid_term <- rbind.fill(silo3_pro_GOid_term,silo3_pro_GOid_term_row)
}

#add protein IDs back to GO IDs
silo3_pro_GOid_term <- cbind(silo3_pro_GO[,"protein_ID"], silo3_pro_GOid_term)
#this results in a table with protein ID listed in one column and each next column contains a GO ID that was listed with the protein in Uniprot DB.


#reshape data so that all GO ID columns are gathered in one column called "GO" 
STACKED_silo3_pro_GOid_term <- tidyr::gather(silo3_pro_GOid_term,"protein_ID","GO", 2:ncol(silo3_pro_GOid_term))
#exlude middle column which just contains the string "protein_ID" in each row
STACKED_silo3_pro_GOid_term <- STACKED_silo3_pro_GOid_term[,c(1,3)]
#remove duplicate rows
STACKED_silo3_pro_GOid_term <- unique(STACKED_silo3_pro_GOid_term)
colnames(STACKED_silo3_pro_GOid_term)[1] <- "protein_ID"
#remove any rows where GO column has NA value. 
STACKED_silo3_pro_GOid_term <- STACKED_silo3_pro_GOid_term[which(!is.na(STACKED_silo3_pro_GOid_term$GO)),]
#this resulting data frame has two columns "protein_ID" and "GO"
```

get GO slim terms for comprehensive proteome from 23C exposured seed
```{r}
 ##get go slim terms for each day
STACKED_silo3_pro_GOid_term_sr_labGOslim <- merge(STACKED_silo3_pro_GOid_term, sr_lab_goslim, by = "GO", all.x = TRUE)
  
STACKED_silo3_pro_GOid_term_sr_labGOslim$GOSlim_bin <- ifelse(is.na(STACKED_silo3_pro_GOid_term_sr_labGOslim$GOSlim_bin),"unmapped",STACKED_silo3_pro_GOid_term_sr_labGOslim$GOSlim_bin)
  
silo3_all_srlab_terms <- data.frame(table(STACKED_silo3_pro_GOid_term_sr_labGOslim$GOSlim_bin))

silo3_all_srlab_BPterms <- data.frame(table(STACKED_silo3_pro_GOid_term_sr_labGOslim[which(STACKED_silo3_pro_GOid_term_sr_labGOslim$aspect=="P"),"GOSlim_bin"]))

```

get GO terms for comprehensive proteome from 29C exposured seed
```{r}
#Subset "protein_ID" column and "GO IDs" column containing list of GO_IDs separated by ";" from all_sig0.1_pro_logFC_pval 
silo9_pro_GO <- silo9_t_nozero_unip_mapped[,c("protein_ID","GO_IDs")]

silo9_pro_GOid_term <- data.frame()
#loop through each line of data frame containing one column with "protein_ID" and one column with list of "GO_IDs" separated by ";"
for (i in 1:nrow(silo9_pro_GO)){
  #create a row that lists all GO IDs associated with one protein listed across rows, where each different GO ID gets put in it's own column
  silo9_pro_GOid_term_row <- data.frame(t(data.frame(strsplit(as.character(silo9_pro_GO$GO_IDs[i]),'; ', fixed = TRUE))), stringsAsFactors = FALSE)
  #add each row created in the line above to the empty data frame
  #this will add NAs in columns where less terms exist; e.g. if a protein only has two terms and another has 10, there will be 8 NAs added to the row for the protein with 2 terms
  silo9_pro_GOid_term <- rbind.fill(silo9_pro_GOid_term,silo9_pro_GOid_term_row)
}

#add protein IDs back to GO IDs
silo9_pro_GOid_term <- cbind(silo9_pro_GO[,"protein_ID"], silo9_pro_GOid_term)
#this results in a table with protein ID listed in one column and each next column contains a GO ID that was listed with the protein in Uniprot DB.


#reshape data so that all GO ID columns are gathered in one column called "GO" 
STACKED_silo9_pro_GOid_term <- tidyr::gather(silo9_pro_GOid_term,"protein_ID","GO", 2:ncol(silo9_pro_GOid_term))
#exlude middle column which just contains the string "protein_ID" in each row
STACKED_silo9_pro_GOid_term <- STACKED_silo9_pro_GOid_term[,c(1,3)]
#remove duplicate rows
STACKED_silo9_pro_GOid_term <- unique(STACKED_silo9_pro_GOid_term)
colnames(STACKED_silo9_pro_GOid_term)[1] <- "protein_ID"
#remove any rows where GO column has NA value. 
STACKED_silo9_pro_GOid_term <- STACKED_silo9_pro_GOid_term[which(!is.na(STACKED_silo9_pro_GOid_term$GO)),]
#this resulting data frame has two columns "protein_ID" and "GO"
```

get GO slim terms for comprehensive proteome from 29C exposured seed
```{r}
 ##get go slim terms for each day
STACKED_silo9_pro_GOid_term_sr_labGOslim <- merge(STACKED_silo9_pro_GOid_term, sr_lab_goslim, by = "GO", all.x = TRUE)
  
STACKED_silo9_pro_GOid_term_sr_labGOslim$GOSlim_bin <- ifelse(is.na(STACKED_silo9_pro_GOid_term_sr_labGOslim$GOSlim_bin),"unmapped",STACKED_silo9_pro_GOid_term_sr_labGOslim$GOSlim_bin)
  
silo9_all_srlab_terms <- data.frame(table(STACKED_silo9_pro_GOid_term_sr_labGOslim$GOSlim_bin))

silo9_all_srlab_BPterms <- data.frame(table(STACKED_silo9_pro_GOid_term_sr_labGOslim[which(STACKED_silo9_pro_GOid_term_sr_labGOslim$aspect=="P"),"GOSlim_bin"]))

```

compare GO slim frequencies between temperatures
```{r}
compare_terms <- merge(silo3_all_srlab_BPterms, silo9_all_srlab_BPterms, by = c("Var1"))
colnames(compare_terms) <- c("GOslim", "23C", "29C")
print(compare_terms)
```

pie charts of 23C and 29C comprehensive proteomes
```{r, pie_charts_Temp_comprehensive_proteome}
ggplot(silo3_all_srlab_BPterms, aes(x = "", y = Freq, fill = Var1)) + geom_bar(stat = "identity") + coord_polar("y", start = 0) + scale_fill_manual(values = getPalette(nrow(silo3_all_srlab_BPterms))) + blank_theme + theme(axis.text.x = element_blank()) + ggtitle("functional characterization of 23C comprehensive proteome")

ggplot(silo9_all_srlab_BPterms, aes(x = "", y = Freq, fill = Var1)) + geom_bar(stat = "identity") + coord_polar("y", start = 0) + scale_fill_manual(values = getPalette(nrow(silo9_all_srlab_BPterms))) + blank_theme + theme(axis.text.x = element_blank()) + ggtitle("functional characterization of 29C comprehensive proteome")
```

##  NSAF


reformat to long list
```{r}
NSAF_t <- data.frame(t(NSAF[,-c(1,2)]))
colnames(NSAF_t) <- paste0("D",NSAF$day,NSAF$temp)
```

add protein column
```{r}
NSAF_t$protein_ID <- rownames(NSAF_t)
NSAF_t <- NSAF_t[grep("CHOYP", NSAF_t$protein_ID),]
```

merge with uniprot
```{r}
#combine uniprot and NSAF data
NSAF_t_unip<- merge(NSAF_t, uniprot, by = "protein_ID", all.x = TRUE)
```

exclude proteins that didn't map to uniprot DB
```{r}
NSAF_t_unip_mapped <- NSAF_t_unip[-grep("unmapped", NSAF_t_unip$Entry),]
nrow(NSAF_t_unip_mapped)
NSAF_t_unip_mapped <- NSAF_t_unip_mapped[which(NSAF_t_unip_mapped$evalue <= 10^-10),]
nrow(NSAF_t_unip_mapped)
```

get GO terms for comprehensive proteome from 23C exposured seed
```{r}
#Subset "protein_ID" column and "GO IDs" column containing list of GO_IDs separated by ";"  
NSAF_pro_GO <- NSAF_t_unip_mapped[,c("protein_ID","GO_IDs")]
head(NSAF_pro_GO)
NSAF_pro_GOid_term <- data.frame()
#loop through each line of data frame containing one column with "protein_ID" and one column with list of "GO_IDs" separated by ";"
for (i in 1:nrow(NSAF_pro_GO)){
  #create a row that lists all GO IDs associated with one protein listed across rows, where each different GO ID gets put in it's own column
  NSAF_pro_GOid_term_row <- data.frame(t(data.frame(strsplit(as.character(NSAF_pro_GO$GO_IDs[i]),'; ', fixed = TRUE))), stringsAsFactors = FALSE)
  #add each row created in the line above to the empty data frame
  #this will add NAs in columns where less terms exist; e.g. if a protein only has two terms and another has 10, there will be 8 NAs added to the row for the protein with 2 terms
  NSAF_pro_GOid_term <- rbind.fill(NSAF_pro_GOid_term,NSAF_pro_GOid_term_row)
}

#add protein IDs back to GO IDs
NSAF_pro_GOid_term <- cbind(NSAF_pro_GO[,"protein_ID"], NSAF_pro_GOid_term)
#this results in a table with protein ID listed in one column and each next column contains a GO ID that was listed with the protein in Uniprot DB.
head(NSAF_pro_GOid_term)

#reshape data so that all GO ID columns are gathered in one column called "GO" 
STACKED_NSAF_pro_GOid_term <- tidyr::gather(NSAF_pro_GOid_term,"protein_ID","GO", 2:ncol(NSAF_pro_GOid_term))
#exlude middle column which just contains the string "protein_ID" in each row
STACKED_NSAF_pro_GOid_term <- STACKED_NSAF_pro_GOid_term[,c(1,3)]
#remove duplicate rows
STACKED_NSAF_pro_GOid_term <- unique(STACKED_NSAF_pro_GOid_term)
colnames(STACKED_NSAF_pro_GOid_term)[1] <- "protein_ID"
#remove any rows where GO column has NA value. 
STACKED_NSAF_pro_GOid_term <- STACKED_NSAF_pro_GOid_term[which(!is.na(STACKED_NSAF_pro_GOid_term$GO)),]
#this resulting data frame has two columns "protein_ID" and "GO"
head(STACKED_NSAF_pro_GOid_term)
```

get GO slim terms for comprehensive proteome from 23C exposured seed
```{r}
 ##get go slim terms for each day
STACKED_NSAF_pro_GOid_term_sr_labGOslim <- merge(STACKED_NSAF_pro_GOid_term, sr_lab_goslim, by = "GO", all.x = TRUE)
  
STACKED_NSAF_pro_GOid_term_sr_labGOslim$GOSlim_bin <- ifelse(is.na(STACKED_NSAF_pro_GOid_term_sr_labGOslim$GOSlim_bin),"unmapped",STACKED_NSAF_pro_GOid_term_sr_labGOslim$GOSlim_bin)
  
NSAF_all_srlab_terms <- data.frame(table(STACKED_NSAF_pro_GOid_term_sr_labGOslim$GOSlim_bin))

NSAF_all_srlab_BPterms <- data.frame(table(STACKED_NSAF_pro_GOid_term_sr_labGOslim[which(STACKED_NSAF_pro_GOid_term_sr_labGOslim$aspect=="P"),"GOSlim_bin"]))

```

pie charts of comprehensive proteome of all time points and temps
```{r, pie_chart_comprehensive_proteome}
ggplot(NSAF_all_srlab_BPterms, aes(x = "", y = Freq, fill = Var1)) + geom_bar(stat = "identity") + coord_polar("y", start = 0) + scale_fill_manual(values = getPalette(nrow(NSAF_all_srlab_BPterms))) + blank_theme + theme(axis.text.x = element_blank()) + ggtitle("functional characterization comprehensive proteome all time points and temps")
```